### 2024-02-13 paper/Untitled.Rmd
update qmd and pdf

---
title: "regression_model"
output: html_document
date: "2024-02-13"
---

library(readxl)
library(readr)
library(haven)
library(tidyverse)

# Reading the datasets
abortion_policy <- read_dta("144981-V1/annual_policy/policyvars01_19.dta")
numbirths <- read_csv("144981-V1/annual_policy/numbirths_2001_2019.csv")
population_data <- read_excel("144981-V1/UKCPR_National_Welfare_Data_Update_020623.xlsx", sheet = "Data")

# Filtering columns and filtering by year
abortion_policy_filtered <- abortion_policy %>%
  select(year, stname, delay, aca_mcaid_exp, pcthsdrop, pcthsgrad, pctsomecol) %>%
  filter(year >= 2001 & year <= 2019)

population_data_filtered <- population_data %>%
  select(state_name, year, 
         Governor_is_Democrat = `Governor is Democrat (1=Yes)`, 
         State_Minimum_Wage = `State Minimum Wage`) %>%
  filter(year >= 2001 & year <= 2019) %>%
  mutate(Governor_is_Democrat = na_if(Governor_is_Democrat, '.'))

numbirths_filtered <- numbirths %>%
  filter(year >= 2001 & year <= 2019)

# Merging data sets and removing rows with NA values
merged_data <- numbirths_filtered %>%
  left_join(abortion_policy_filtered, by = c("year", "stname" = "stname")) %>%
  left_join(population_data_filtered, by = c("stname" = "state_name", "year")) %>%
  drop_na(Governor_is_Democrat, delay, aca_mcaid_exp, pcthsdrop, pcthsgrad, pctsomecol, numbirth1544, State_Minimum_Wage)

# Running the regression model on the cleaned and filtered dataset
model <- lm(numbirth1544 ~ delay + aca_mcaid_exp + pcthsdrop + pcthsgrad + pctsomecol + Governor_is_Democrat + State_Minimum_Wage, data = merged_data)

# Summary of the model
summary(model)

###### Checking assumptions ######
install.packages("car")
install.packages("lmtest")
install.packages("sandwich")

library(tidyverse)
library(car)
library(lmtest)
library(sandwich)

# Check for linearity and homoscedasticity
plot(model, which = 1:2)

# Check for normality of residuals with a Q-Q plot
qqPlot(model, main = "Q-Q Plot")

# Test for independence of residuals (Durbin-Watson test)
dwtest(model)

# Check for multicollinearity with Variance Inflation Factors (VIF)
vif(model) # A VIF value > 5 or 10 indicates high multicollinearity

# Check for homoscedasticity with Breusch-Pagan test
bptest(model)

# Check for non-random residuals
# Residuals vs. Fitted values plot
plot(model, which = 1)
abline(h = 0, col = "red")