### 2024-02-13 paper/paper.qmd
changes to pdf and scripts
Added graphs to paper output and changed the path for various scripts

---
title: "Red States = Red Flags?"
subtitle: "An analysis of birth rate trends from 2001 to 2019 and an exploration of political affiliation of different states and Birth rates in the United States,"
author: 
  - Rayan Awad Alim
  - Maria Mangru
  - Isha Junega
thanks: "Code and data are available at: https://github.com/RayanAlim/Birth_Rate_Analysis"
date: today
date-format: long
abstract: "First sentence. Second sentence. Third sentence. Fourth sentence."
format: pdf
number-sections: true
bibliography: references.bib
editor: 
  markdown: 
    wrap: 72
---

```{r echo=FALSE, message=FALSE}
#| include: false
#| warning: false
#| message: false

library(tidyverse)

```

Introduction

broader context to motivate The birth rate is one of the most important
determinants of a country’s population growth, potentially impacting
policy decisions about the health system, education, and economy (1).
https://www.waldenu.edu/programs/health/resource/why-birth-rates-are-significant-in-health-studies-and-health-science

some detail about what the paper is about This paper dives into the
potential factors that may cause a birth rate to decrease. a clear gap
that needs to be filled Not sure what was done We replicated the paper
by Kearney, Levine, andPardue (2022) and focused on wages, access to
healthcare (with religious importance) and childcare expenditure data to
draw conclusions about whether these factors affect the decline or birth
rates what was found

why is it important the structure of the paper

Data

Article:

Model

Model Information Dependent variable: birth rate per 1000 population
Independent variables:State min wage, Abortion delay, Child enforcement
expenditure, Political affiliation

Education

Model set-up

Model defined as:

```{=tex}
\begin{align} 
y_i|\mu_i, \sigma &\sim \mbox{Normal}(\mu_i, \sigma) \\
\mu_i &= \alpha + \beta_i + \gamma_i\\
\alpha &\sim \mbox{Normal}(0, 2.5) \\
\beta &\sim \mbox{Normal}(0, 2.5) \\
\gamma &\sim \mbox{Normal}(0, 2.5) \\
\sigma &\sim \mbox{Exponential}(1)
\end{align}
```
We run the model in R [@citeR] using the rstanarm package of @rstanarm.
We use the default priors from rstanarm.

Model justification

We expect a positive relationship between the size of the wings and time
spent aloft. In particular...

We can use maths by including latex between dollar signs, for instance
$\theta$.

Results

Our results are summarized in @tbl-modelresults.

```{r echo=FALSE, message=FALSE}
#| echo: false
#| eval: true
#| warning: false
#| message: false

library(rstanarm)

first_model <-
  readRDS(file = here::here("models/first_model.rds"))

```

1- Medicaid coverage has increased which allows more low-income people
to get an abortion.

2- Maximum well-fare has gone down so it more difficult to raise a kid

3- Parents seem to have more of an interest in having an abortion which
aligns with the decrease in number of births

{r} #\| echo: false #\| eval: true #\| label: tbl-modelresults #\|
tbl-cap: "Explanatory models of flight time based on wing width and wing
length" #\| warning: false

modelsummary::modelsummary( list( "First model" = first_model ),
statistic = "mad", fmt = 2 )

Analysis Results

### Analysis Results

# State-Specific Birth Averages (2001-2019):

The average number of births varies considerably from state to state,
with California (CA) having the highest average number of births and
Vermont (VT) having the lowest.

Larger states, both in terms of geography and population like Texas (TX)
and New York (NY), show significantly higher average numbers of births,
which aligns with the expected outcomes based on population.

Smaller states such as Vermont (VT) and North Dakota (ND) have lower
average numbers of births, which may reflect not just smaller
populations but also demographic trends such as aging populations.

```{r echo=FALSE, message=FALSE}
library(haven)
library(tidyverse)
library(tidyr)

# Importing the CSV files
numbirths_2001_2019 <- read_csv("../data/analysis_data/annual_policy/numbirths_2001_2019.csv")
births_educ_race_age6 <- read_csv("../data/analysis_data/decomp/births_educ_race_age6.csv")
numbirths_educ_2044 <- read_csv("../data/analysis_data/educ/numbirths_educ_2044.csv")
state_births_04_19 <- read_csv("../data/analysis_data/map/state_births_04_19.csv")

# Display the first few rows of each data-frame to verify the imports
view(numbirths_2001_2019)
view(births_educ_race_age6)
view(numbirths_educ_2044)
view(state_births_04_19)

### Number of Births ### 
colnames(numbirths_2001_2019)

average_births_by_state <- numbirths_2001_2019 %>%
  filter(year >= 2001, year <= 2019) %>%
  group_by(stname) %>%
  summarize(AverageBirths = mean(numbirth1544, na.rm = TRUE)) %>%
  ungroup()
view(average_births_by_state)

ggplot(average_births_by_state, aes(y = reorder(stname, AverageBirths), x = AverageBirths)) +
  geom_bar(stat = "identity", fill = "steelblue") +
  theme_minimal() +
  labs(title = "Average Number of Births by State (2001-2019)",
       y = "State",
       x = "Average Number of Births") +
  theme(axis.text.y = element_text(angle = 0)) 
```

# State Minimum Wage Changes (2001-2019):

The District of Columbia (DC) exhibits the largest increase, with its
minimum wage rising by \$7.85, from \$6.15 in 2001 to \$14.00 in 2019.
Arizona (AZ) and Colorado (CO) also show substantial increases of \$6.85
each.

Georgia (GA) and Wyoming (WY) stand out as the only states with no
change in minimum wage, maintaining a rate of \$5.15 across the years.
It is worth noting that this rate is equal to the federal minimum wage
between 1997 - 2007, suggesting that these states have not adjusted
their minimum wage above the federal floor set at that time.

```{r echo=FALSE, message=FALSE, results='asis'}
library(knitr)
library(readxl)
library(dplyr)
library(ggplot2)

# Load Data 
welfare_data <- read_excel("../data/UKCPR_National_Welfare_Data_Update_020623.xlsx", sheet = "Data")

# Retrieving data for state minimum wage for 2001 - 2019
filtered_minwage <- welfare_data %>%
  select(state_name, year, `State Minimum Wage`) %>%
  filter(year >= 2001, year <= 2019)

##### State Minimum Wage #####
# Filter data for 2001 and 2019
minwage_2001 <- welfare_data %>%
  filter(year == 2001) %>%
  select(state_name, year, `State Minimum Wage`)

minwage_2019 <- welfare_data %>%
  filter(year == 2019) %>%
  select(state_name, year, `State Minimum Wage`)

# Rename the `State Minimum Wage` columns to reflect the year
minwage_2001 <- minwage_2001 %>%
  rename(MinWage2001 = `State Minimum Wage`)

minwage_2019 <- minwage_2019 %>%
  rename(MinWage2019 = `State Minimum Wage`)

# Merge the 2001 and 2019 data by state_name
minwage_change <- left_join(minwage_2001, minwage_2019, by = "state_name")

# Calculate the change in minimum wage from 2001 to 2019
minwage_change <- minwage_change %>%
  mutate(ChangeInMinWage = MinWage2019 - MinWage2001) %>%
  select(state_name, MinWage2001, MinWage2019, ChangeInMinWage)

kable(minwage_change, format = "markdown", caption = "Change in State Minimum Wage from 2001 to 2019")

```

# State-Specific Births by Abortion Delay Policy from 2001 to 2019:

The trend lines across the graphs indicate a general decline in birth
rates. The decline is more pronounced in states that have implemented
delay policies, suggesting a possible correlation between policy
enactment and a faster reduction in birth rates.

States with Delay

Birth rates vary significantly between states. For example, Arkansas
(AR) shows a sharp decrease in birth rates compared to other states. In
contrast, Pennsylvania (PA) has a gentler slope.

The range of birth rates in 2001 spans from approximately 17 per 1000
(excluding Utah (UT)) people in certain states to below 15 per 1000 in
others, with a convergence of rates around 12 to 14 per 1000.

The states in this graph are ones which had a policy change from no
delay to delay which allows for the analysis of the before-and-after
effects of policy implementation.

For example, Arizona's (AZ) implementation year of 2009 shows birth rate
descends sharply afterward. In contrast, North Carolina (NC) in 2011
didn't show such an immediate shift.

The states without delay policies showcase a natural decline in birth
rates over time, without the potential influence of policy changes. This
trend is consistent across the states, with some variations in the
steepness of the decline.

For example, California (CA) maintains a relatively high birth rate over
the years, starting near 15 per 1000 in 2001 and ending around 12 per
1000 in 2019. This could be influenced by its socio-economic profile and
immigration patterns.

All states show a general decline in relative birth rates over the
19-year period, which is in line with the national trend of declining
birth rates.

State-Specific Trends:

California (CA): The trend in California shows a relatively steady
decline throughout the period, without any sharp changes.

Florida (FL): Florida's trend is similar to California's, with a gradual
decline.

North Carolina (NC): North Carolina, with a policy change in 201, shows
a relatively stable trend before the policy change and a sharper decline
after.

Texas (TX): Texas exhibited a noticeable decline after the policy
implementation year of 2003.

When contrasting the birth rates of Texas and California, two states
with large populations, we see that Texas's birth rates declined more
sharply after its policy change in 2003 compared to California. This
might suggest an association between the policy implementation and the
accelerated decline in birth rates in Texas.

<!-- -->

Comparing Georgia and North Carolina, both of which implemented delay
policies. Georgia's birth rates were rising before the policy change and
then started declining after, while North Carolina's birth rates were
relatively stable before 2011 and then declined.

Discussion

## First discussion point: Abortion

#relative average \# of births in states delay

```{r echo=FALSE, message=FALSE}
library(readxl)
library(dplyr)
library(readr)

# Determine which states had a delay or no delay policy 
abortion_policy <- read_dta("../data/analysis_data/annual_policy/policyvars01_19.dta") %>%
  filter(year >= 2001, year <= 2019) %>%
  select(delay, stname, year)

# Load the population data from the Excel file
population_data <- read_excel("../data/UKCPR_National_Welfare_Data_Update_020623.xlsx", sheet = "Data") %>%
  select(state_name, year, Population) %>%
  filter(year >= 2001, year <= 2019)

numbirths_data <- read_csv("../data/analysis_data/annual_policy/numbirths_2001_2019.csv") %>%
  rename(births = numbirth1544) %>%
  filter(year >= 2001, year <= 2019)

# Ensure both datasets have the 'stname' column for joining
population_data <- population_data %>%
  rename(stname = state_name)

# Merge the number of births data with the population data
births_population_data <- numbirths_data %>%
  left_join(population_data, by = c("year", "stname"))

# Calculate the relative average number of births per 1000 people
births_population_data <- births_population_data %>%
  mutate(relative_births = (births / Population) * 1000)

# Load the abortion policy data
abortion_policy <- read_dta("../data/analysis_data/annual_policy/policyvars01_19.dta") %>%
  select(year, stname, delay)

# Merge the relative births data with the abortion policy data
combined_data <- abortion_policy %>%
  left_join(births_population_data, by = c("year", "stname"))

# Define the groups based on predefined 'states_no_delay' and 'states_with_delay'
states_no_delay <- c("AK", "AL", "CA", "CO", "CT", "DC", "DE", "FL", "HI", "IA", "IL", "MA", "MD", "ME", "MT", "NH", "NJ", "NM", "NV", "NY", "OR", "RI", "TN", "VT", "WA", "WY")
states_with_delay <- c("AR", "ID", "IN", "KS", "KY", "LA", "MI", "MO", "MS", "ND", "NE", "OH", "PA", "SC", "SD", "UT", "VA", "WI")

# Filter the combined_data for states with no delay
combined_data_no_delay <- combined_data %>%
  filter(stname %in% states_no_delay)

# Filter the combined_data for states with delay
combined_data_with_delay <- combined_data %>%
  filter(stname %in% states_with_delay)

# Plot for states with policy delay
ggplot(data = combined_data_with_delay) +
  geom_line(aes(x = year, y = relative_births, group = stname, color = stname)) +
  labs(title = "Relative Average Number of Births in States With Delay (2001-2019)",
       x = "Year",
       y = "Relative Average Number of Births per 1000 People",
       color = "State") +
  theme_minimal()

```

#relative average \# of births in states w/o delay

```{r echo=FALSE, message=FALSE}
library(readxl)
library(dplyr)
library(readr)

# Determine which states had a delay or no delay policy 
abortion_policy <- read_dta("../data/analysis_data/annual_policy/policyvars01_19.dta") %>%
  filter(year >= 2001, year <= 2019) %>%
  select(delay, stname, year)

# Load the population data from the Excel file
population_data <- read_excel("../data/UKCPR_National_Welfare_Data_Update_020623.xlsx", sheet = "Data") %>%
  select(state_name, year, Population) %>%
  filter(year >= 2001, year <= 2019)

numbirths_data <- read_csv("../data/analysis_data/annual_policy/numbirths_2001_2019.csv") %>%
  rename(births = numbirth1544) %>%
  filter(year >= 2001, year <= 2019)

# Ensure both datasets have the 'stname' column for joining
population_data <- population_data %>%
  rename(stname = state_name)

# Merge the number of births data with the population data
births_population_data <- numbirths_data %>%
  left_join(population_data, by = c("year", "stname"))

# Calculate the relative average number of births per 1000 people
births_population_data <- births_population_data %>%
  mutate(relative_births = (births / Population) * 1000)

# Load the abortion policy data
abortion_policy <- read_dta("../data/analysis_data/annual_policy/policyvars01_19.dta") %>%
  select(year, stname, delay)

# Merge the relative births data with the abortion policy data
combined_data <- abortion_policy %>%
  left_join(births_population_data, by = c("year", "stname"))

# Define the groups based on predefined 'states_no_delay' and 'states_with_delay'
states_no_delay <- c("AK", "AL", "CA", "CO", "CT", "DC", "DE", "FL", "HI", "IA", "IL", "MA", "MD", "ME", "MT", "NH", "NJ", "NM", "NV", "NY", "OR", "RI", "TN", "VT", "WA", "WY")
states_with_delay <- c("AR", "ID", "IN", "KS", "KY", "LA", "MI", "MO", "MS", "ND", "NE", "OH", "PA", "SC", "SD", "UT", "VA", "WI")

# Filter the combined_data for states with no delay
combined_data_no_delay <- combined_data %>%
  filter(stname %in% states_no_delay)

# Filter the combined_data for states with delay
combined_data_with_delay <- combined_data %>%
  filter(stname %in% states_with_delay)

# Plot for states with no policy delay
ggplot(data = combined_data_no_delay) +
  geom_line(aes(x = year, y = relative_births, group = stname, color = stname)) +
  labs(title = "Relative Average Number of Births in States Without Delay (2001-2019)",
       x = "Year",
       y = "Relative Average Number of Births per 1000 People",
       color = "State") +
  theme_minimal()

```

#relative birth rates in transition states

```{r echo=FALSE, message=FALSE}
library(readxl)
library(dplyr)
library(readr)

# Determine which states had a delay or no delay policy 
abortion_policy <- read_dta("../data/analysis_data/annual_policy/policyvars01_19.dta") %>%
  filter(year >= 2001, year <= 2019) %>%
  select(delay, stname, year)

# Load the population data from the Excel file
population_data <- read_excel("../data/UKCPR_National_Welfare_Data_Update_020623.xlsx", sheet = "Data") %>%
  select(state_name, year, Population) %>%
  filter(year >= 2001, year <= 2019)

numbirths_data <- read_csv("../data/analysis_data/annual_policy/numbirths_2001_2019.csv") %>%
  rename(births = numbirth1544) %>%
  filter(year >= 2001, year <= 2019)

# Ensure both datasets have the 'stname' column for joining
population_data <- population_data %>%
  rename(stname = state_name)

# Merge the number of births data with the population data
births_population_data <- numbirths_data %>%
  left_join(population_data, by = c("year", "stname"))

# Calculate the relative average number of births per 1000 people
births_population_data <- births_population_data %>%
  mutate(relative_births = (births / Population) * 1000)

# Load the abortion policy data
abortion_policy <- read_dta("../data/analysis_data/annual_policy/policyvars01_19.dta") %>%
  select(year, stname, delay)

# Merge the relative births data with the abortion policy data
combined_data <- abortion_policy %>%
  left_join(births_population_data, by = c("year", "stname"))

# Define the groups based on predefined 'states_no_delay' and 'states_with_delay'
states_no_delay <- c("AK", "AL", "CA", "CO", "CT", "DC", "DE", "FL", "HI", "IA", "IL", "MA", "MD", "ME", "MT", "NH", "NJ", "NM", "NV", "NY", "OR", "RI", "TN", "VT", "WA", "WY")
states_with_delay <- c("AR", "ID", "IN", "KS", "KY", "LA", "MI", "MO", "MS", "ND", "NE", "OH", "PA", "SC", "SD", "UT", "VA", "WI")

# Filter the combined_data for states with no delay
combined_data_no_delay <- combined_data %>%
  filter(stname %in% states_no_delay)

# Filter the combined_data for states with delay
combined_data_with_delay <- combined_data %>%
  filter(stname %in% states_with_delay)

# States and years of policy implementation
policy_changes <- data.frame(
  stname = c("AZ", "GA", "MN", "NC", "OK", "TX", "WV"),
  year_implemented = c(2009, 2005, 2003, 2011, 2005, 2003, 2003)
)

# Filter combined_data
transition_states_data <- combined_data %>%
  filter(stname %in% policy_changes$stname)

# Creating the graph
ggplot(data = transition_states_data) +
  geom_line(aes(x = year, y = relative_births, group = stname, color = stname)) +
  geom_vline(data = policy_changes, aes(xintercept = year_implemented, color = stname), linetype="dashed") +
  geom_text(data = policy_changes, aes(x = year_implemented, y = Inf, label = year_implemented), vjust = -0.5, color = "black") +
  scale_x_continuous(breaks = seq(2001, 2019, by = 2)) +
  labs(title = "Transition to Policy Delay and Relative Average Number of Births (2001-2019)",
       x = "Year",
       y = "Relative Average Number of Births per 1000 People",
       color = "State") +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))

```

#relative birth rates in selected states

```{r echo=FALSE, message=FALSE}
library(readxl)
library(dplyr)
library(readr)

# Determine which states had a delay or no delay policy 
abortion_policy <- read_dta("../data/analysis_data/annual_policy/policyvars01_19.dta") %>%
  filter(year >= 2001, year <= 2019) %>%
  select(delay, stname, year)

# Load the population data from the Excel file
population_data <- read_excel("../data/UKCPR_National_Welfare_Data_Update_020623.xlsx", sheet = "Data") %>%
  select(state_name, year, Population) %>%
  filter(year >= 2001, year <= 2019)

numbirths_data <- read_csv("../data/analysis_data/annual_policy/numbirths_2001_2019.csv") %>%
  rename(births = numbirth1544) %>%
  filter(year >= 2001, year <= 2019)

# Ensure both datasets have the 'stname' column for joining
population_data <- population_data %>%
  rename(stname = state_name)

# Merge the number of births data with the population data
births_population_data <- numbirths_data %>%
  left_join(population_data, by = c("year", "stname"))

# Calculate the relative average number of births per 1000 people
births_population_data <- births_population_data %>%
  mutate(relative_births = (births / Population) * 1000)

# Load the abortion policy data
abortion_policy <- read_dta("../data/analysis_data/annual_policy/policyvars01_19.dta") %>%
  select(year, stname, delay)

# Merge the relative births data with the abortion policy data
combined_data <- abortion_policy %>%
  left_join(births_population_data, by = c("year", "stname"))

# Define the groups based on predefined 'states_no_delay' and 'states_with_delay'
states_no_delay <- c("AK", "AL", "CA", "CO", "CT", "DC", "DE", "FL", "HI", "IA", "IL", "MA", "MD", "ME", "MT", "NH", "NJ", "NM", "NV", "NY", "OR", "RI", "TN", "VT", "WA", "WY")
states_with_delay <- c("AR", "ID", "IN", "KS", "KY", "LA", "MI", "MO", "MS", "ND", "NE", "OH", "PA", "SC", "SD", "UT", "VA", "WI")

# Filter the combined_data for states with no delay
combined_data_no_delay <- combined_data %>%
  filter(stname %in% states_no_delay)

# Filter the combined_data for states with delay
combined_data_with_delay <- combined_data %>%
  filter(stname %in% states_with_delay)

# Step 1: Select states for comparison
states_comparison <- combined_data %>%
  filter(stname %in% c("CA", "TX", "GA", "NC", "FL"))

# Step 2: Calculate descriptive statistics and visualize trends
# For California (no policy change) and Texas (policy change in 2003)
ca_tx_data <- states_comparison %>%
  filter(stname %in% c("CA", "TX"))

# Calculate summary statistics for CA and TX before and after 2003
ca_tx_summary <- ca_tx_data %>%
  group_by(stname) %>%
  summarise(
    mean_before = mean(relative_births[year < 2003], na.rm = TRUE),
    mean_after = mean(relative_births[year >= 2003], na.rm = TRUE)
  )

# For Georgia (policy change in 2005), North Carolina (policy change in 2011), and Florida (no policy change)
ga_nc_fl_data <- states_comparison %>%
  filter(stname %in% c("GA", "NC", "FL"))

# Calculate summary statistics for GA and NC before and after policy changes, and compare with FL
ga_nc_fl_summary <- ga_nc_fl_data %>%
  group_by(stname) %>%
  summarise(
    mean_before = mean(relative_births[year < if_else(stname == "GA", 2005, if_else(stname == "NC", 2011, year))], na.rm = TRUE),
    mean_after = mean(relative_births[year >= if_else(stname == "GA", 2005, if_else(stname == "NC", 2011, year))], na.rm = TRUE)
  )

# Step 3: Visualize trends
ggplot(states_comparison, aes(x = year, y = relative_births, color = stname)) +
  geom_line() +
  geom_vline(xintercept = c(2003, 2005, 2011), linetype = "dashed", color = "black") +
  labs(title = "Relative Birth Rates in Selected States (2001-2019)", y = "Relative Births per 1000 People", x = "Year") +
  theme_minimal()

```

Second discussion: Political affiliation

{r} \# POLITICAL AFFILIATION MAP

Third discussion point: Minimum Wage

Fourth discussion point: Education

{r} #Education graph

Weaknesses and next steps

Weaknesses and next steps should also be included.

\newpage

\appendix

Appendix

Additional data details

Model details

Posterior predictive check

In @fig-ppcheckandposteriorvsprior-1 we implement a posterior predictive
check. This shows...

In @fig-ppcheckandposteriorvsprior-2 we compare the posterior with the
prior. This shows...

{r} #\| eval: true #\| echo: false #\| message: false #\| warning: false
#\| label: fig-ppcheckandposteriorvsprior #\| layout-ncol: 2 #\|
fig-cap: "Examining how the model fits, and is affected by, the data"
#\| fig-subcap: \["Posterior prediction check", "Comparing the posterior
with the prior"\]

pp_check(first_model) + theme_classic() + theme(legend.position =
"bottom")

posterior_vs_prior(first_model) + theme_minimal() +
scale_color_brewer(palette = "Set1") + theme(legend.position =
"bottom") + coord_flip()

Diagnostics

@fig-stanareyouokay-1 is a trace plot. It shows... This suggests...

@fig-stanareyouokay-2 is a Rhat plot. It shows... This suggests...

{r} #\| echo: false #\| eval: true #\| message: false #\| warning: false
#\| label: fig-stanareyouokay #\| fig-cap: "Checking the convergence of
the MCMC algorithm" #\| fig-subcap: \["Trace plot", "Rhat plot"\] #\|
layout-ncol: 2

plot(first_model, "trace")

plot(first_model, "rhat")

\newpage

References