### 2024-02-13 child_expendure_graphs.qmd

---
title: "graph"
format: html
editor: visual
---

## Quarto

Quarto enables you to weave together content and executable code into a finished document. To learn more about Quarto see <https://quarto.org>.

## Set Up

Setting up the library and getting the data.

```{r}
#### Workspace setup ####
library(tidyverse)
library(dplyr)
library(haven)
library(stata2r)

#### Download data ####

data <- read_dta("data/long_term/childcare_asec_09_1519.dta")
head(data)
```

You can add options to executable code like this

```{r}
group

ggplot(data, aes(x=year, y=obs)) +
    geom_col() 
```

```{r}
#### Download data ####

child_enforce_data <- read_dta("data/annual_policy/policyvars01_19.dta")
head(child_enforce_data)

# medicaid benefits
ggplot(child_enforce_data, aes(x=year, y=aca_mcaid_exp)) +
    geom_col()

# max well fair benefits
ggplot(child_enforce_data, aes(x=year, y=logben3)) +
    geom_point()

#parental consent for abortion
#ggplot(child_enforce_data, aes(x = year, y = parental)) +
#  geom_bar(stat = "identity")

#ggplot(child_enforce_data, aes(x = parental)) +
#  geom_bar(position = "dodge")

# Bar plot
ggplot(child_enforce_data, aes(x = factor(year), fill = factor(parental))) +
  geom_bar(position = "stack") +
  labs(x = "Year", y = "Count", fill = "Interest", title = "Parental Abortion Intrest Distribution Over the Years")
```