### 2024-02-15 paper/paper.qmd

---
title: "Red States = Red Flags?"
subtitle: "An analysis of birth rate trends from 2001 to 2019 and an exploration of political affiliation of different states and Birth rates in the United States,"
author: 
  - Rayan Awad Alim
  - Maria Mangru
  - Isha Junega
thanks: "Code and data are available at: https://github.com/RayanAlim/Birth_Rate_Analysis"
date: today
date-format: long
abstract: "First sentence. Second sentence. Third sentence. Fourth sentence."
format: pdf
number-sections: true
#bibliography: references.bib
editor: 
  markdown: 
    wrap: 72
---

```{r echo=FALSE, message=FALSE}
#| include: false
#| warning: false
#| message: false

### Preparing Workspace
# Install and load necessary packages
if (!requireNamespace("tidyverse", quietly = TRUE)) install.packages("tidyverse")
if (!requireNamespace("dplyr", quietly = TRUE)) install.packages("dplyr")
if (!requireNamespace("tidyr", quietly = TRUE)) install.packages("tidyr")
if (!requireNamespace("ggplot2", quietly = TRUE)) install.packages("ggplot2")
if (!requireNamespace("knitr", quietly = TRUE)) install.packages("knitr")
if (!requireNamespace("readr", quietly = TRUE)) install.packages("readr")
if (!requireNamespace("maps", quietly = TRUE)) install.packages("maps")
if (!requireNamespace("haven", quietly = TRUE)) install.packages("haven")
if (!requireNamespace("readxl", quietly = TRUE)) install.packages("readxl")
if (!requireNamespace("here", quietly = TRUE)) install.packages("here")
if (!requireNamespace("lmtest", quietly = TRUE)) install.packages("lmtest")
if (!requireNamespace("sandwich", quietly = TRUE)) install.packages("sandwich")

# Workspace Set up
library(tidyverse)
library(dplyr)
library(tidyr)
library(ggplot2)
library(knitr)
library(readr)
library(maps)
library(haven)
library(readxl)
library(here)
library(lmtest)
library(sandwich)
```

Introduction

A country's birth rate is a key determinant of its population growth,
which has a profound impact on policy pertaining to education, health,
and the economy in general. With changing societal expectations and
family dynamics, birth rates across developed countries have seemingly
been declining in recent decades. The replacement birth rate is 2.1
births per female, and significant deviations from this figure have
consequences that affect countries in several important ways. A high
birth rate means that society has to adapt to support a growing
population, which places a stress on resources like food and housing.
However, a low birth rate means that a society cannot sustain their
population and elders may be viewed as “burdens”. The contributing
factors that drive birth rates vary depending by country. This paper
will focus on exploring findings related to declining birth rates in the
United States.

Melissa S. Kearney, Phillip B. Levine, and Luke Pardue, analyze
potential reasons for a decline in birth rate in their paper, “The
Puzzle of Falling US Birth Rates since the Great Recession.” In their
analysis, examine birth rates by demographic groups defined by age,
education, race and ethnicity, marital status, and birth parity inorder
to find clues that serve as an explanation for the decline. They found
factors pointing to the time and location do not serve as enough
explanation to the decline in birthrates, although less quantifiable,
women shifting their priorities to their aspirations for life instead of
having more children is an important contributor. Their analysis does
not explore solid evidence of US-specific policies or economic factors
that can explain the depth of that decline(). We address this gap by
replicating their paper and focusing on wages, access to healthcare
(with religious importance) and childcare expenditure data to draw
conclusions about whether these factors affect the decline or birth
rates. We found that XX. This is important because. This paper is
structured to include the data, results, and the discussion that include
our analysis. Although the code used in the original paper is Strata, we
use R to replicate some of its graphs, as well as to conduct analysis of
our own.

Data

Melissa S. Kearney, Phillip B. Levine, and Luke Pardue's paper "The
Puzzle of Falling US Birth Rates since the Great Recession" provided one
of the datasets used in this analysis, notably the policyvars01_19
dataset. This dataset presents a thorough summary of state-level
policies that may have an impact on birth rates in the US from 2001 to
2019 by assembling data on numerous policy variables from numerous
sources.

Article:

Model

As part of the investigation, the researchers investigated the impact of
the abortion policy delay on relative number of births per 1000 people
using a Difference-in-Differences approach. This was done to compare the
change in outcomes over time for the policy between a treatment group
and a control group. In this paper, we looked at the effect of the
abortion delay policy implementation in North Carolina in 2011. The
control group of choice was Tennesse as this state did not implement an
abortion delay policy during the years 2001 - 2019. Tennesse was also
selected as the control group due to its geographical proximity to North
Carolina and demographic similarities. This was done to satisfy the
parallel trends assumption. The model employed is as follows:

$[ \text{Relative Birthsit} = \beta_0 + \beta_1 (\text{Treatment}_i) + \beta_2(\text{PostPolicy}_t) + \beta_3(\text{Treatment} \times \text{PostPolicy}t) + \varepsilon{it} ]$

Model set-up
=======
Where:

-   Relative Births~it~ is the number of births per 1000 people in state
    i at time t

-   $Treatment_i$ is an indicator for weather state i is NC

-   $PostPolicy_t$ is an indicator for years after the policy change

-   $Treatment_i$ x $PostPolicy_t$ is the interaction term

Results

1- Medicaid coverage has increased which allows more low-income people
to get an abortion.

1- Medicaid coverage has increased which allows more low-income people
to get an abortion.

2- Maximum well-fare has gone down so it more difficult to raise a kid

3- Parents seem to have more of an interest in having an abortion which
aligns with the decrease in number of births

2- Maximum well-fare has gone down so it more difficult to raise a kid

3- Parents seem to have more of an interest in having an abortion which
aligns with the decrease in number of births

Analysis Results

### Analysis Results

# State-Specific Birth Averages (2001-2019):

Figure 1 presents a summary of the average number of scaled births by
state over the years 2001 to 2109. Generally, it can be seen that
certain states exhibit higher average number of births, while others,
usually smaller states, exhibit lower number of births. The three states
with the highest averages of births are California (CA), Texas (TX) and
New York (NY).

```{r echo=FALSE, message=FALSE}

# Importing the CSV file
numbirths_2001_2019 <- read_csv("../data/analysis_data/annual_policy/numbirths_2001_2019.csv")

# Calculate average births scaled down by 1000 for visual representation
average_births_by_state <- numbirths_2001_2019 %>%
  filter(year >= 2001, year <= 2019) %>%
  group_by(stname) %>%
  summarize(AverageBirths = mean(numbirth1544, na.rm = TRUE) / 1000) %>% # Scale down by a factor of 1000
  ungroup()

# Create the bar plot with scaled averages
ggplot(average_births_by_state, aes(x = reorder(stname, -AverageBirths), y = AverageBirths)) +
  geom_bar(stat = "identity", fill = "steelblue") +
  theme_minimal() +
  labs(title = "Scaled Average Number of Births by State (2001-2019)",
       x = "State",
       y = "Scaled Average Number of Births") +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))
```

# Child Expenditure

Figure C2 The number of people with ACA Medicaid increased dramatically
from 2000 to about 2014. Then, since 2014, it has slowly increased
non-decreasingly till 2020.

Figure C3 There has been a linear decrease of the maximum welfare
benefits from 2000 to 2015. From 2015 to 2020 there has been a small
rise in the maximum welfare benefits. Overall, there has been a loss in
the maximum welfare benefits, despite the small rise.

Figure C4 The blue portion of each column represents the number of
parents who have consent for an abortion. The red portion represents the
number of parents who don’t have consent for an abortion. There is a
slight increase in consent for abortion from 2001 to 2019

# State Minimum Wage Changes (2001-2019):

Figure 5 shows the trend of the average minimum wage across states
depicted by the blue line and the average number of births represented
by the red line, between the years 2001 to 2009. The trend for average
minimum wage shows a noticeable increase over time. In contrast, the
average number of births shows a fluctuating trend. Starting from 2001,
the number of births gradually increase, with a peak occurring around
2007. After this peak, there is a discernible decline in the average
number of births.

```{r echo=FALSE, message=FALSE, results='asis'}

# Load the data sets
welfare_data <- read_excel("../data/UKCPR_National_Welfare_Data_Update_020623.xlsx", sheet = "Data") %>%
  select(state_name, year, `State Minimum Wage`) %>%
  filter(year >= 2001 & year <= 2019)

numbirths_2001_2019 <- read_csv("../data/analysis_data/annual_policy/numbirths_2001_2019.csv") %>%
  select(year, stname, numbirth1544) %>%
  filter(year >= 2001 & year <= 2019) %>%
  rename(state_name = stname)

# Merge data sets on state_name and year
merged_data <- merge(welfare_data, numbirths_2001_2019, by = c("state_name", "year"))

# Calculate average minimum wage and average number of births per year 
average_data <- merged_data %>%
  group_by(year) %>%
  summarise(Average_Minimum_Wage = mean(`State Minimum Wage`, na.rm = TRUE),
            Average_Births = mean(numbirth1544, na.rm = TRUE))

# Find maximum values for scaling 
max_births <- max(average_data$Average_Births, na.rm = TRUE)
max_wage <- max(average_data$Average_Minimum_Wage, na.rm = TRUE)
scaling_factor <- max_wage / max_births

# Plot the graph with adjusted scales
ggplot(average_data) +
  geom_line(aes(x = year, y = Average_Minimum_Wage), colour = "blue", size = 1) +
  geom_line(aes(x = year, y = Average_Births * scaling_factor), colour = "red", size = 1) +
  scale_y_continuous(
    "Minimum Wage",
    sec.axis = sec_axis(~ . / scaling_factor, name = "Number of Births")
  ) +
  theme_minimal() +
  labs(title = "Comparison of Average Minimum Wage and Birth Rates (2001-2009)",
       x = "Year")

```

# Impact of Abortion Delay Policy

Abortion delay policies refer to state-level regulations which stipulate
a required waiting period between a woman’s initial consultation for an
abortion and the procedure itself. This paper looks at states which have
consistently implemented an abortion delay policy, those without this
policy and those which have shifted from having no delay to have a delay
policy between the years 2001 to 2019.

```{r echo=FALSE, message=FALSE, results='asis'}
# Define file paths 
policy_data_path <- "../data/analysis_data/annual_policy/policyvars01_19.dta"
population_data_path <- "../data/UKCPR_National_Welfare_Data_Update_020623.xlsx"
births_data_path <- "../data/analysis_data/annual_policy/numbirths_2001_2019.csv"
```

The range of birth rates in 2001 spans from approximately 17 per 1000
(excluding Utah (UT)) people in certain states to below 15 per 1000 in
others, with a convergence of rates around 12 to 14 per 1000.

The states in this graph are ones which had a policy change from no
delay to delay which allows for the analysis of the before-and-after
effects of policy implementation.

For example, Arizona's (AZ) implementation year of 2009 shows birth rate
descends sharply afterward. In contrast, North Carolina (NC) in 2011
didn't show such an immediate shift.

The states without delay policies showcase a natural decline in birth
rates over time, without the potential influence of policy changes. This
trend is consistent across the states, with some variations in the
steepness of the decline.

For example, California (CA) maintains a relatively high birth rate over
the years, starting near 15 per 1000 in 2001 and ending around 12 per
1000 in 2019. This could be influenced by its socio-economic profile and
immigration patterns.

All states show a general decline in relative birth rates over the
19-year period, which is in line with the national trend of declining
birth rates.

State-Specific Trends:

California (CA): The trend in California shows a relatively steady
decline throughout the period, without any sharp changes.

Florida (FL): Florida's trend is similar to California's, with a gradual
decline.

North Carolina (NC): North Carolina, with a policy change in 201, shows
a relatively stable trend before the policy change and a sharper decline
after.

Texas (TX): Texas exhibited a noticeable decline after the policy
implementation year of 2003.

When contrasting the birth rates of Texas and California, two states
with large populations, we see that Texas's birth rates declined more
sharply after its policy change in 2003 compared to California. This
might suggest an association between the policy implementation and the
accelerated decline in birth rates in Texas.

Comparing Georgia and North Carolina, both of which implemented delay
policies. Georgia's birth rates were rising before the policy change and
then started declining after, while North Carolina's birth rates were
relatively stable before 2011 and then declined.

# Education

```{r echo=FALSE, message=FALSE}
#Education

# Load Education data
educ_data <- read_csv("../data/analysis_data/educ/numbirths_educ_2044.csv")

# Creating the plot
ggplot(educ_data, aes(x = year)) +
  geom_line(aes(y = numbirth_lesshs, color = "Less than HS"), linewidth = 1) +
  geom_line(aes(y = numbirth_hsgrad, color = "HS Graduate"), linewidth = 1) +
  geom_line(aes(y = numbirth_somecoll, color = "Some College"), linewidth = 1) +
  geom_line(aes(y = numbirth_colgrad, color = "College Graduate"), linewidth = 1) +
  geom_line(
    aes(y = numbirthalleduc, color = "All Education Levels"),
    size = 1,
    linetype = "dashed"
  ) +
  scale_color_manual(
    values = c(
      "Less than HS" = "blue",
      "HS Graduate" = "green",
      "Some College" = "orange",
      "College Graduate" = "red",
      "All Education Levels" = "grey"
      )
  ) +
  labs(
    title = "Number of Births by Education Level Over Time",
    x = "Year",
    y = "Number of Births",
    color = "Education Level"
  ) +
  theme_minimal() +
  theme(legend.title = element_text(size = 12),
        legend.text = element_text(size = 10))

# Display the plot
ggsave("births_by_education_level_over_time.png",
       width = 10,
       height = 6)

```

# State Political Affiliation and Birth Rates
```{r echo=FALSE, message=FALSE}
# https://stackoverflow.com/questions/29381652/plotting-us-heat-map-with-text

#reading the UKNCPR Datat
poldata <- read_csv(here("data", "analysis_data", "UKCPR_National_Welfare_Data_Update_020623.csv"))

poldata <- poldata %>%
  select(state_name,
         year,
         Population,
         `Fraction of State Senate that is Democrat`)

# Create a mapping of state abbreviations to full names
state_abbrevs <- state.abb
state_names <- state.name

# Create a data frame for the mapping
state_mapping <-
  data.frame(abbrev = tolower(state_abbrevs),
             name = tolower(state_names))

#Get numbirths data
numbirths_data <- read_csv(here("data", "analysis_data", "annual_policy", "numbirths_2001_2019.csv"))
births_2019 <- numbirths_data %>% filter(year == 2019)

# Convert state abbreviations to full names in your datasets
numbirths_2019_fullname <- births_2019 %>%
  mutate(stname = tolower(stname)) %>%
  left_join(state_mapping, by = c("stname" = "abbrev")) %>%
  select(-stname) %>%
  rename(stname = name)

poldata_2019 <- poldata %>% filter(year == 2019)

poldata_2019_fullname <- poldata_2019 %>%
  mutate(state_name = tolower(state_name)) %>%
  left_join(state_mapping, by = c("state_name" = "abbrev")) %>%
  select(-state_name) %>%
  rename(state_name = name)

# Merge poldata with numbirths_data using full state names
data_merged <-
  left_join(numbirths_2019_fullname,
            poldata_2019_fullname,
            by = c("stname" = "state_name"))

states_map <- map_data("state")

# Merge with states map data
map_data_merged <-
  left_join(states_map, data_merged, by = c("region" = "stname"))

# Plot
ggplot(map_data_merged) +
  geom_polygon(
    aes(
      x = long,
      y = lat,
      group = group,
      fill = `Fraction of State Senate that is Democrat`
    ),
    color = "white"
  ) +
  geom_text(
    data = map_data_merged %>% group_by(region) %>% summarise(
      long = mean(long),
      lat = mean(lat),
      numbirths = first(numbirth1544)
    ),
    aes(x = long, y = lat, label = numbirths),
    size = 3,
    check_overlap = TRUE
  ) +
  scale_fill_gradient(low = "red",
                      high = "blue",
                      name = "Democrat Fraction") +
  coord_fixed(1.3) +
  labs(title = "2019 Births and Democrat Fraction in State Senate by State") +
  theme_void() +
  theme(legend.position = "right")
```

Discussion

## First discussion point: Abortion

#relative average \# of births in states delay

```{r echo=FALSE, message=FALSE}
# Determine which states had a delay or no delay policy 
abortion_policy <- read_dta("../data/analysis_data/annual_policy/policyvars01_19.dta") %>%

# Load and prepare the abortion policy data
abortion_policy <- read_dta(policy_data_path) %>%
  filter(year >= 2001, year <= 2019) %>%
  select(year, stname, delay)

# Load and prepare the population data
population_data <- read_excel(population_data_path, sheet = "Data") %>%
  select(state_name, year, Population) %>%
  filter(year >= 2001, year <= 2019) %>%
  rename(stname = state_name)

# Load and prepare the number of births data
numbirths_data <- read_csv(births_data_path) %>%
  rename(births = numbirth1544) %>%
  filter(year >= 2001, year <= 2019)

# Merge the datasets
births_population_data <- numbirths_data %>%
  left_join(population_data, by = c("year", "stname")) %>%
  mutate(relative_births = (births / Population) * 1000)

combined_data <- abortion_policy %>%
  left_join(births_population_data, by = c("year", "stname"))

# Define the groups based on predefined 'states_no_delay' and 'states_with_delay'
states_no_delay <- c("AK", "AL", "CA", "CO", "CT", "DC", "DE", "FL", "HI", "IA", "IL", "MA", "MD", "ME", "MT", "NH", "NJ", "NM", "NV", "NY", "OR", "RI", "TN", "VT", "WA", "WY")
states_with_delay <- c("AR", "ID", "IN", "KS", "KY", "LA", "MI", "MO", "MS", "ND", "NE", "OH", "PA", "SC", "SD", "UT", "VA", "WI")

# Filter data for states with no delay and plot
combined_data_no_delay <- combined_data %>%
  filter(stname %in% states_no_delay)

ggplot(data = combined_data_no_delay, aes(x = year, y = relative_births, group = stname, color = stname)) +
  geom_line() +
  labs(title = "Relative Average Number of Births in States Without Delay (2001-2019)",
       x = "Year",
       y = "Relative Average Number of Births per 1000 People",
       color = "State") +
  theme_minimal()
```

Across the 19-year span, the data for states which have not implemented
an abortion policy showcases a wide range of trends. Several states
demonstrate a peak in birth rates at various points during the study
period. While there is variation between the states, the overall
trajectory indicated that the early to late 2000s were a period of
higher birth rates, with a shift toward lower birth rates post 2010. It
is important to note that this data in Figure 7 showcases the relative
average number of births per 1000 people and fluctuates from 10 to
approximately 16.

```{r echo=FALSE, message=FALSE, results='asis'}
# Define file paths 
policy_data_path <- "../data/analysis_data/annual_policy/policyvars01_19.dta"
population_data_path <- "../data/UKCPR_National_Welfare_Data_Update_020623.xlsx"
births_data_path <- "../data/analysis_data/annual_policy/numbirths_2001_2019.csv"

# Load and prepare the abortion policy data
abortion_policy <- read_dta(policy_data_path) %>%
  filter(year >= 2001, year <= 2019) %>%
  select(year, stname, delay)

# Load and prepare the population data
population_data <- read_excel(population_data_path, sheet = "Data") %>%
  select(state_name, year, Population) %>%
  filter(year >= 2001, year <= 2019) %>%
  rename(stname = state_name)

# Load and prepare the number of births data
numbirths_data <- read_csv(births_data_path) %>%
  rename(births = numbirth1544) %>%
  filter(year >= 2001, year <= 2019)

# Merge the datasets
births_population_data <- numbirths_data %>%
  left_join(population_data, by = c("year", "stname")) %>%
  mutate(relative_births = (births / Population) * 1000)

combined_data <- abortion_policy %>%
  left_join(births_population_data, by = c("year", "stname"))

# Define the groups based on predefined 'states_no_delay' and 'states_with_delay'
states_no_delay <- c("AK", "AL", "CA", "CO", "CT", "DC", "DE", "FL", "HI", "IA", "IL", "MA", "MD", "ME", "MT", "NH", "NJ", "NM", "NV", "NY", "OR", "RI", "TN", "VT", "WA", "WY")
states_with_delay <- c("AR", "ID", "IN", "KS", "KY", "LA", "MI", "MO", "MS", "ND", "NE", "OH", "PA", "SC", "SD", "UT", "VA", "WI")

# Filter data for states with delay and plot
combined_data_with_delay <- combined_data %>%
  filter(stname %in% states_with_delay)

ggplot(data = combined_data_with_delay, aes(x = year, y = relative_births, group = stname, color = stname)) +
  geom_line() +
  labs(title = "Relative Average Number of Births in States With Delay (2001-2019)",
       x = "Year",
       y = "Relative Average Number of Births per 1000 People",
       color = "State") +
  theme_minimal()
```

In contrast, figure \[\] looks at the trend of relative average number
of births which have consistently implemented a delay policy. Similar to
states which have not implemented a delay policy, birth rates for these
states have generally decreased over the 19 year period. Over this
period, the relative average number of births per 1000 people fluctuated
significantly, ranging from around 10 to over 20.

```{r echo=FALSE, message=FALSE, results='asis'}

# Define file paths 
policy_data_path <- "../data/analysis_data/annual_policy/policyvars01_19.dta"
population_data_path <- "../data/UKCPR_National_Welfare_Data_Update_020623.xlsx"
births_data_path <- "../data/analysis_data/annual_policy/numbirths_2001_2019.csv"

# Load and prepare the abortion policy data
abortion_policy <- read_dta(policy_data_path) %>%
  filter(year >= 2001, year <= 2019) %>%
  select(year, stname, delay)

# Load and prepare the population data
population_data <- read_excel(population_data_path, sheet = "Data") %>%
  select(state_name, year, Population) %>%
  filter(year >= 2001, year <= 2019) %>%
  rename(stname = state_name)

# Load and prepare the number of births data
numbirths_data <- read_csv(births_data_path) %>%
  rename(births = numbirth1544) %>%
  filter(year >= 2001, year <= 2019)

# Merge the datasets
births_population_data <- numbirths_data %>%
  left_join(population_data, by = c("year", "stname")) %>%
  mutate(relative_births = (births / Population) * 1000)

combined_data <- abortion_policy %>%
  left_join(births_population_data, by = c("year", "stname"))

# Define the groups based on predefined 'states_no_delay' and 'states_with_delay'
states_no_delay <- c("AK", "AL", "CA", "CO", "CT", "DC", "DE", "FL", "HI", "IA", "IL", "MA", "MD", "ME", "MT", "NH", "NJ", "NM", "NV", "NY", "OR", "RI", "TN", "VT", "WA", "WY")
states_with_delay <- c("AR", "ID", "IN", "KS", "KY", "LA", "MI", "MO", "MS", "ND", "NE", "OH", "PA", "SC", "SD", "UT", "VA", "WI")

# States which had a change in policy and plot
policy_changes <- data.frame(
  stname = c("AZ", "GA", "MN", "NC", "OK", "TX", "WV"),
  year_implemented = c(2009, 2005, 2003, 2011, 2005, 2003, 2003)
)

transition_states_data <- combined_data %>%
  filter(stname %in% policy_changes$stname)

# Filter and plot for states which had a change in policy
transition_states_data <- combined_data %>%
  filter(stname %in% policy_changes$stname)

ggplot(data = transition_states_data, aes(x = year, y = relative_births, group = stname, color = stname)) +
  geom_line() +
  geom_vline(data = policy_changes, aes(xintercept = year_implemented, color = stname), linetype="dashed") +
  geom_text(data = policy_changes, aes(x = year_implemented, y = Inf, label = year_implemented), vjust = -0.5, color = "black") +
  scale_x_continuous(breaks = seq(2001, 2019, by = 2)) +
                       labs(title = "Transition to Policy Delay and Relative Average Number of Births (2001-2019)",
                            x = "Year",
                            y = "Relative Average Number of Births per 1000 People",
                            color = "State") +
                       theme_minimal() +
                       theme(axis.text.x = element_text(angle = 45, hjust = 1))
```

Lastly, figure \[\] looks at states which transitioned from having no
abortion delay to policy to implementing this policy at various years
(indicated by the broken lines). A common trend among these states was
an initial period of relatively higher birth rates in the early 2000s,
with subsequent declines over the years. This trend also closely follows
what observed for states which implemented an abortion delay policy and
those which have not implemented such a policy.

# Difference-in-Difference

```{r echo=FALSE, message=FALSE, results='asis'}
# Define file paths 
policy_data_path <- "../data/analysis_data/annual_policy/policyvars01_19.dta"
population_data_path <- "../data/UKCPR_National_Welfare_Data_Update_020623.xlsx"
births_data_path <- "../data/analysis_data/annual_policy/numbirths_2001_2019.csv"

# Load and prepare the abortion policy data
abortion_policy <- read_dta(policy_data_path) %>%
  filter(year >= 2001, year <= 2019) %>%
  select(year, stname, delay)

# Load and prepare the population data
population_data <- read_excel(population_data_path, sheet = "Data") %>%
  select(state_name, year, Population) %>%
  filter(year >= 2001, year <= 2019) %>%
  rename(stname = state_name)

# Load and prepare the number of births data
numbirths_data <- read_csv(births_data_path) %>%
  rename(births = numbirth1544) %>%
  filter(year >= 2001, year <= 2019)

# Merge the datasets
births_population_data <- numbirths_data %>%
  left_join(population_data, by = c("year", "stname")) %>%
  mutate(relative_births = (births / Population) * 1000)

combined_data <- abortion_policy %>%
  left_join(births_population_data, by = c("year", "stname"))

# Difference-in-Difference Analysis for NC with a policy change in 2011 compared to a no-delay state
did_data <- combined_data %>%
  filter(stname %in% c("NC", "TN"), year >= 2001, year <= 2019) %>%
  mutate(treatment = if_else(stname == "NC", 1, 0),
         post_policy = if_else(year >= 2011, 1, 0),
         treatment_x_post = treatment * post_policy)

# Fit the DiD model
did_model <- lm(relative_births ~ treatment + post_policy + treatment_x_post, data = did_data)

# Summary of the DiD model and robust standard errors
summary_did <- summary(did_model)
robust_summary <- coeftest(did_model, vcov = vcovHC(did_model, type = "HC1"))

# Extract coefficients and their standard errors
coefficients <- coef(robust_summary)
std_errors <- sqrt(diag(vcovHC(did_model, type = "HC1")))

# Combine coefficients and standard errors into a data frame
coef_df <- data.frame(Estimate = coefficients, `Std. Error` = std_errors)

# Output the robust summary as a table
kable(coef_df, format = "markdown")

```

These abortion delay policies created natural experiments which can be
used to apply a Difference-in-Differences approach. The results from
comparing North Carolina (treatment) with Tennessee (control) are as
follow: - The intercept ($B_0$) estimate is 13.4345, suggesting that the
baseline relative number of births per 1000 people is 13.4345 in the
absence of treatment and before the policy change. - The treatment
effect ($B_1$) is estimated at 0.4887, which is statistically
significant (p = 0.01078). This indicates that, prior to the policy
change, NC had 0.4887 more births per 1000 people than TN. - The
post_policy effect ($B_2$) is -1.2372, also statistically significant (p
= 1.26e-07), suggesting a general decrease in relative births per 1000
people after 2011 in both states. - The interaction term ($B_3$), which
represents the estimated effect of the policy change, is -0.7770, with a
p-value of 0.00569, indicating that the policy delay in NC is associated
with a reduction of 0.7770 births per 1000 people relative to TN after
the policy implementation. - The model explains a significant portion of
the variance in relative births per 1000 people ($R\^2$ = 0.8267),
indicating a good fit.

# Education

#Politicial Affiliation

### Discussion

# Medicaid - remove heading after The Affordable Care
(ACA)’s Medicaid is “intended to address systemic health inequalities
for millions of Americans who lack health insurance” Medicaid is
“expected to provide coverage to low-income individuals,” and also
covers low-income pregnant women. Excluding the states where abortion is
banned, Medicaid covers abortions, making it more accessible. According
to Figure C1, More people continue to have medicaid as seen based on the
trend from 2000 to 2020, making an affordable abortion more accessible.
In Figure C3, we can see that there is an increase in parental consent
for abortion. In other words, more parents are increasingly interested
in getting an abortion every year. A potential motivator for parents to
want an abortion is that as a low-income family, it is harder to take
care of a larger family. Welfare is “federal government-sponsored
assistance programs for individuals and families in need.” In Figure C2,
it is evident that there is a decrease in maximum welfare benefits,
making it less affordable for lower-income families to have a larger
family size.The increasing demand for an abortion and the insurance
covering it may contribute to a decline in birthrates.

# Abortion policy - remove heading after

Through the difference-in-difference analysis, the results indicate a
decrease of approximately 0.777 births per 1000 people in North Carolina
post policy implementation relative to TN, which did not implement the
abortion delay policy. This decrease is statistically significant,
suggesting that implementing an abortion delay policy may have a real
effect on birth rates.

The observed decrease in birth rates in North Carolina post 2011 could
be partially explained by increased barriers to access as highlighted by
de Londras et al. whose review supports the results derived in this
paper. This review summarizes the effects of mandated waiting periods
(MWPs) on outcomes connected to abortion. Their analysis suggests that
MWPs may cause delays in obtaining abortion services, which may lower
the number of births particularly in young, unmarried women. Noting the
disproportionate impact on marginalized populations, such as teenagers,
persons of color, and those with lower socioeconomic standing, they also
draw attention to the broader human rights concerns of MWPs.

The results of the DiD found in this paper are supported by strong
standard errors. However, we need to consider the limitations of the DiD
approach as there are several other factors which may impact birth rates
such as economic fluctuations and access to healthcare. Future
iterations of this paper could expand on this research by conducting
multi-state analysis for a variety of confounding factors.

Second discussion: Political affiliation and birth rates

From our analysis of political affiliations over the years, there are some relation between the two, especially in years following aggresive rhetorics or policies about reproductive rights. This indicates that periods following aggressive political rhetoric or significant policy changes regarding reproductive rights often coincide with noticeable changes in birth rates. This trend is observable across various states and is particularly pronounced in areas where political affiliation (# of democrats in state senate) change very differently from one side to the other. The rhetoric employed by political leaders and the nature of the policies they implement can have a substantial impact on individuals' choices regarding family planning. For instance, policies that limit access to reproductive healthcare, including contraception and abortion services, can lead to an increase in birth rates, especially in regions where alternative options are scarce.[@Nargund2009DecliningBirthRate]

Conversely, political movements that advocate for and expand access to reproductive healthcare are often associated with a stabilization or decrease in birth rates. This can be attributed to increased autonomy over reproductive choices, allowing individuals to make informed decisions about if and when to have children.

Fourth discussion point: Education

```{r}
#Education graph
```

Weaknesses and next steps A weakness of this paper is a lack of
economical context. For example, we analyze our data independent of
events like the Recession that happened in 2008. Without providing
context to current events in the time period, our paper gives a weaker
justification for the decline in birthrates. Next steps for improving
our analysis include not only providing context for the events happening
in the US, but focusing on the birth rates on a more micro level. Since
each state within the US has its own laws, it will provide more insight
to focus on the birth rates within each state. Additionally, to improve
our analysis furthermore, we can compare the US to other countries that
are similar from an economical standpoint and compare their birth rate
to the US.

Weaknesses and next steps should also be included.

\newpage

\appendix

Appendix

Additional data details

Model details

```{r}

```