### 2024-02-15 paper/paper.qmd
submission

---
title: "The Decline of US Birth Rates: Analyzing Minimum Wages, Abortion Policies, Education, and Child Support Enforcement"
subtitle:  "Reproduction of ‘The Puzzle of Falling US Birth Rates since the Great Recession’ (Kearney, Levine and, Pardue,2022)"

author: 
  - Rayan Awad Alim
  - Maria Mangru
  - Isha Juneja
thanks: "Code and data are available at: https://github.com/RayanAlim/Birth_Rate_Analysis"
date: today
date-format: long
abstract: "Melissa S. Kearney, Phillip B. Levine, and Luke Pardue in their paper, 
'The Puzzle of Falling US Birth Rates since the Great Recession', analyze the 
decreasing birth rates within the US. In this paper, we repicate use their data 
and replicate their results of their study. While they focused on examining 
birth rates by demographics, we focus on minimum wages, education, abortion 
poilicy, and child support enforcement. The importance of our research is that 
it highlights the underlying cause of why parents may not to have many children, 
which leads to a decline in birthrates."
format: pdf
number-sections: true
#bibliography: references.bib
editor: 
  markdown: 
    wrap: 72
---

```{r echo=FALSE, message=FALSE}
#| include: false
#| warning: false
#| message: false

### Preparing Workspace
# Install and load necessary packages
if (!requireNamespace("tidyverse", quietly = TRUE)) install.packages("tidyverse")
if (!requireNamespace("dplyr", quietly = TRUE)) install.packages("dplyr")
if (!requireNamespace("tidyr", quietly = TRUE)) install.packages("tidyr")
if (!requireNamespace("ggplot2", quietly = TRUE)) install.packages("ggplot2")
if (!requireNamespace("knitr", quietly = TRUE)) install.packages("knitr")
if (!requireNamespace("readr", quietly = TRUE)) install.packages("readr")
if (!requireNamespace("maps", quietly = TRUE)) install.packages("maps")
if (!requireNamespace("haven", quietly = TRUE)) install.packages("haven")
if (!requireNamespace("readxl", quietly = TRUE)) install.packages("readxl")
if (!requireNamespace("here", quietly = TRUE)) install.packages("here")
if (!requireNamespace("lmtest", quietly = TRUE)) install.packages("lmtest")
if (!requireNamespace("sandwich", quietly = TRUE)) install.packages("sandwich")

# Workspace Set up
library(tidyverse)
library(dplyr)
library(tidyr)
library(ggplot2)
library(knitr)
library(readr)
library(maps)
library(haven)
library(readxl)
library(here)
library(lmtest)
library(sandwich)

# manages graph placement
knitr::opts_chunk$set(fig.pos = "H", out.extra = "")
```

# Introduction {#sec-intro}

A country's birth rate is a key determinant of its population growth,
which has a profound impact on policy pertaining to education, health,
and the economy in general. With changing societal expectations and
family dynamics, birth rates across developed countries have seemingly
been declining in recent decades. The replacement birth rate is 2.1
births per female, and significant deviations from this figure have
consequences that affect countries in several important ways. A high
birth rate means that society has to adapt to support a growing
population, which places a stress on resources like food and housing.
However, a low birth rate means that a society cannot sustain their
population and elders may be viewed as “burdens”. The contributing
factors that drive birth rates vary depending by country. This paper
will focus on exploring findings related to declining birth rates in the
United States.

Melissa S. Kearney, Phillip B. Levine, and Luke Pardue, analyze
potential reasons for a decline in birth rate in their paper, “The
Puzzle of Falling US Birth Rates since the Great Recession.” In their
analysis, examine birth rates by demographic groups defined by age,
education, race and ethnicity, marital status, and birth parity in order
to find clues that serve as an explanation for the decline. They found
factors pointing to the time and location do not serve as enough
explanation to the decline in birthrates, although less quantifiable,
women shifting their priorities to their aspirations for life instead of
having more children is an important contributor. Their analysis does
not explore solid evidence of US-specific policies or economic factors
that can explain the depth of that decline. We address this gap by
replicating their paper and focusing on wages, abortion policies,
education and childcare support enforcement data to draw conclusions
about whether these factors affect the decline or birth rates.

In this paper, we created a linear regression model to determine the
impact of abortion policy on number of birthrates. We analyzed the
affects of education on the number of birthrates by tracking the the
birthrate for each education level separately. Additionally, to gage the
impact of childcare support enforcement on the number of birthrates, we
looked at how the maximum welfare benefits, including Medicaid, affect
low-income individuals. Analyzing the wages, abortion policies,
education, and childcare support important because they give deeper
insights regarding why parents may not prefer to have many children and
therefore may cause a decline in birthrates. The paper is structured to
include the data, results, and the discussion that include our analysis.
Although the code used in the original paper is Strata, we use R[@citeR]
to replicate some of its graphs, as well as to conduct analysis of our
own.

# Data {#sec-data}

## Data Source and Methodology

As part of our study, we looked closely at the variables influencing the
decline in US birth rates. Our goal was to expand on and close the gaps
left by Melissa S. Kearney, Phillip B. Levine, and Luke Pardue in their
paper, "The Puzzle of Falling US Birth Rates since the Great Recession."
The interaction between birth rates and childcare costs, healthcare
availability, education, political affiliation, and incomes was the main
focus of our research. We replicated many of the visuals and analyses
from the original paper using R for data visualization and analysis,
building on its conclusions through the use of a variety of datasets and
approaches. The primary datasets for our study came from a variety of
sources, such as the UK Center for Poverty Research National Welfare
Database, the annual policy analysis data, and comprehensive birth
records from 2001 to 2019. The Center for Poverty Research National
Welfare Database [@ukcprWelfareData] at the University of Kentucky
supplied more data on state-level laws, minimum wage levels, and
political affiliations to supplement these original statistics. The
UKCPR National Welfare Data which is a comprehensive dataset that has
been compiled and annually updated by the University of Kentucky Center
for Poverty Research (UKCPR). This database includes many variables such
as population, wage levels, political landscape, dynamics of poverty,
employment, and welfare in the United States at the state level.
Covering a wide time span from 1980 to 2021, this dataset provides a
longitudinal view of significant socio-economic indicators across the
country. Building on earlier work by Kearney and Levine
[@Kearney2022Puzzle], state-specific abortion delay law data was updated
using the Guttmacher Institute's "Counseling and Waiting Periods for
Abortion" database. The Kaiser Family Foundation's "Status of State
Medicaid Expansion Decisions: Interactive Map" (KFF 2021) and a
collection of birth statistics were the sources of information on
Medicaid expansions. The methods used to gather the data made it
possible to analyze state-level policies that potentially affect US
birth rates in detail between 2001 and 2019. Through the integration of
publicly accessible datasets and meticulously curated policy variables,
our study offered a comprehensive understanding of the diverse factors
influencing the fluctuations in US birth rates.

## Variables and Summary Statisitcs

With an emphasis on the years 2001 to 2019, we created and improved a
collection of variables which are needed for studying the decrease in
birth rates throughout the United States. Our results are stronger
because we were able to record the short-term impacts of various
different socioeconomic determinants on birth rates during this time.
The development of important variables, their importance, storage
formats, and the interactions between them are described in detail
below. Graphs, summary statistics, and comments are used to provide
context.

The first main variable of interest is the delay variable which looks at
the Abortion Policy Delay. This binary variable indicated whether or not
a stated has implemented a delay in abortion policy. Our
Difference-in-Differences (DiD) analysis relied on this variable to
evaluate the policy's effect on birth rates. Secondly relative_births
was created to determine the relative birth rates of each state by
population in order to provide a common metric for comparing birth rates
among states with different population sizes. This variable was created
by combining birth statistics with population estimates. State Minimum
Wage (min_wage) indicates the minimum wage for each state in a certain
year. The purpose of including this economic variable in the study of
the correlation between birth rates and salary levels was to test the
hypothesis that family planning choices could be influenced by greater
salaries. Healthcare Access (healthcare_access): An integer-based
composite index based on the availability of healthcare facilities and
the expansion of Medicaid. This variable sought to measure the ease of
access to healthcare services, as this could have an effect on birth
rates by affecting the availability of prenatal and postnatal care.
Next, the variable childcare_exp quantified whether or not parents
received any support for their childcare expenses. Considering the
substantial expenses linked to childrearing, this variable was essential
to comprehending the financial influences on decisions about family
size.

High-level cleaning included resolving missing numbers, guaranteeing
consistency across datasets, and filtering data to the years 2001-2019.
Harmonizing variables such as state_name made proper merging possible.
To aid in analysis, numerical variables were inspected for anomalies
and, if required, normalized.

Visual insights into the effects of abortion policy delays and economic
factors on birth rates were provided via graphs showing the trend in
relative birth rates over time, by state, and by policy status. A
summary of the distribution and dispersion of the data was given via
summary statistics. For example, the distribution of min_wage brought
attention to economic inequities, while the average relative birth rate
and its standard deviation showed variances between states and years.

The associations between the important variables were shown using
scatter plots and correlation analysis. For example, the relationship
between min salary and relative births was investigated in order to
comprehend the impact of the economy on family planning choices.

## Measurements

These datasets have several limitations even if they are quite useful
for our research. Firstly, reporting variations in birth rates can occur
between states and over time. These disparities could be the result of
delayed data entry, modifications to reporting requirements, or
variations in data collection techniques. These discrepancies could
distort the birth rate estimates, causing an overestimation or
underestimating in particular states.

Secondly, although the policy data is extensive, it might not fully
account for all state-level programs that have the potential to affect
birth rates. In addition, they do not give insight into variations
within the policy. Birth trends may also be significantly influenced by
policies that are not covered by the dataset, such as regional
healthcare access initiatives or educational initiatives, but these
factors are not taken into consideration in our research. To address
these limitations, we used several strategies. To confirm the trends
found and guarantee consistency across sources, we cross-referenced our
primary data sources with supplementary datasets and research. Next, in
order to determine the potential impact of data mistakes, we performed
sensitivity analysis. We evaluated the robustness of our findings to
differences in data quality by correcting for any under- or over
reporting in birth data.

# Results: {#sec-results}

## Model

As part of the investigation, the researchers investigated the impact of
the abortion policy delay on relative number of births per 1000 people
using a Difference-in-Differences approach. This was done to compare the
change in outcomes over time for the policy between a treatment group
and a control group. In this paper, we looked at the effect of the
abortion delay policy implementation in North Carolina in 2011. The
control group of choice was Tennesse as this state did not implement an
abortion delay policy during the years 2001 - 2019. Tennesse was also
selected as the control group due to its geographical proximity to North
Carolina and demographic similarities. This was done to satisfy the
parallel trends assumption. The model employed is as follows:

$[ \text{Relative Birthsit} = \beta_0 + \beta_1 (\text{Treatment}_i) + \beta_2(\text{PostPolicy}_t) + \beta_3(\text{Treatment} \times \text{PostPolicy}t) + \varepsilon{it} ]$

### Model set-up

Where:

-   Relative Births~it~ is the number of births per 1000 people in state
    i at time t

-   $Treatment_i$ is an indicator for weather state i is NC

-   $PostPolicy_t$ is an indicator for years after the policy change

-   $Treatment_i$ x $PostPolicy_t$ is the interaction term

## Birth Rate Results

### State-Specific Birth Averages (2001-2019):

Figure 1 presents a summary of the average number of scaled births by
state over the years 2001 to 2109. Generally, it can be seen that
certain states exhibit higher average number of births, while others,
usually smaller states, exhibit lower number of births. The three states
with the highest averages of births are California (CA), Texas (TX) and
New York (NY).

```{r echo=FALSE, message=FALSE}
#| label: fig-1

# Importing the CSV file
numbirths_2001_2019 <- read_csv("../data/analysis_data/annual_policy/numbirths_2001_2019.csv")

# Calculate average births scaled down by 1000 for visual representation
average_births_by_state <- numbirths_2001_2019 %>%
  filter(year >= 2001, year <= 2019) %>%
  group_by(stname) %>%
  summarize(AverageBirths = mean(numbirth1544, na.rm = TRUE) / 1000) %>% # Scale down by a factor of 1000
  ungroup()

# Create the bar plot with scaled averages
ggplot(average_births_by_state, aes(x = reorder(stname, -AverageBirths), y = AverageBirths)) +
  geom_bar(stat = "identity", fill = "steelblue") +
  theme_minimal() +
  labs(title = "Scaled Average Number of Births by State (2001-2019)",
       x = "State",
       y = "Scaled Average Number of Births") +
  theme(axis.text.x = element_text(angle = 90, hjust = 1, size=7))
```

## State Minimum Wage Changes (2001-2019):

Figure 2 shows the trend of the average minimum wage across states
depicted by the blue line and the average number of births represented
by the red line, between the years 2001 to 2009. The trend for average
minimum wage shows a noticeable increase over time. In contrast, the
average number of births shows a fluctuating trend. Starting from 2001,
the number of births gradually increase, with a peak occurring around
2007. After this peak, there is a discernible decline in the average
number of births.

```{r echo=FALSE, message=FALSE, results='asis'}
#| label: fig-2

# Load the data sets
welfare_data <- read_excel("../data/UKCPR_National_Welfare_Data_Update_020623.xlsx", sheet = "Data") %>%
  select(state_name, year, `State Minimum Wage`) %>%
  filter(year >= 2001 & year <= 2019)

numbirths_2001_2019 <- read_csv("../data/analysis_data/annual_policy/numbirths_2001_2019.csv") %>%
  select(year, stname, numbirth1544) %>%
  filter(year >= 2001 & year <= 2019) %>%
  rename(state_name = stname)

# Merge data sets on state_name and year
merged_data <- merge(welfare_data, numbirths_2001_2019, by = c("state_name", "year"))

# Calculate average minimum wage and average number of births per year 
average_data <- merged_data %>%
  group_by(year) %>%
  summarise(Average_Minimum_Wage = mean(`State Minimum Wage`, na.rm = TRUE),
            Average_Births = mean(numbirth1544, na.rm = TRUE))

# Find maximum values for scaling 
max_births <- max(average_data$Average_Births, na.rm = TRUE)
max_wage <- max(average_data$Average_Minimum_Wage, na.rm = TRUE)
scaling_factor <- max_wage / max_births

# Plot the graph with adjusted scales
ggplot(average_data, aes(x = year)) +
  geom_line(aes(y = Average_Minimum_Wage), colour = "blue", linewidth = 1) +
  geom_line(aes(y = Average_Births * scaling_factor), colour = "red", linewidth = 1) +
  scale_y_continuous(
    "Minimum Wage",
    sec.axis = sec_axis(~ . / scaling_factor, name = "Number of Births")
  ) +
  theme_minimal() +
  labs(title = "Comparison of Average Minimum Wage and Birth Rates (2001-2019)",
       x = "Year") +
  theme(plot.title = element_text(size = 10))

```

## Impact of Abortion Delay Policy

Abortion delay policies refer to state-level regulations which stipulate
a required waiting period between a woman’s initial consultation for an
abortion and the procedure itself. This paper looks at states which have
consistently implemented an abortion delay policy, those without this
policy and those which have shifted from having no delay to have a delay
policy between the years 2001 to 2019.

```{r echo=FALSE, message=FALSE, results='asis'}
#| label: fig-3
# Define file paths 
policy_data_path <- "../data/analysis_data/annual_policy/policyvars01_19.dta"
population_data_path <- "../data/UKCPR_National_Welfare_Data_Update_020623.xlsx"
births_data_path <- "../data/analysis_data/annual_policy/numbirths_2001_2019.csv"

# Load and prepare the abortion policy data
abortion_policy <- read_dta(policy_data_path) %>%
  filter(year >= 2001, year <= 2019) %>%
  select(year, stname, delay)

# Load and prepare the population data
population_data <- read_excel(population_data_path, sheet = "Data") %>%
  select(state_name, year, Population) %>%
  filter(year >= 2001, year <= 2019) %>%
  rename(stname = state_name)

# Load and prepare the number of births data
numbirths_data <- read_csv(births_data_path) %>%
  rename(births = numbirth1544) %>%
  filter(year >= 2001, year <= 2019)

# Merge the datasets
births_population_data <- numbirths_data %>%
  left_join(population_data, by = c("year", "stname")) %>%
  mutate(relative_births = (births / Population) * 1000)

combined_data <- abortion_policy %>%
  left_join(births_population_data, by = c("year", "stname"))

# Define the groups based on predefined 'states_no_delay' and 'states_with_delay'
states_no_delay <- c("AK", "AL", "CA", "CO", "CT", "DC", "DE", "FL", "HI", "IA", "IL", "MA", "MD", "ME", "MT", "NH", "NJ", "NM", "NV", "NY", "OR", "RI", "TN", "VT", "WA", "WY")
states_with_delay <- c("AR", "ID", "IN", "KS", "KY", "LA", "MI", "MO", "MS", "ND", "NE", "OH", "PA", "SC", "SD", "UT", "VA", "WI")

# Filter data for states with no delay and plot
combined_data_no_delay <- combined_data %>%
  filter(stname %in% states_no_delay)

ggplot(data = combined_data_no_delay, aes(x = year, y = relative_births, group = stname, color = stname)) +
  geom_line() +
  labs(title = "Relative Average Number of Births in States Without Delay (2001-2019)",
       x = "Year",
       y = "Relative Average Number of Births per 1000 People") +
  theme_minimal() +
  theme(
    plot.title = element_text(size = 8, hjust = 0.5), 
    axis.title.y = element_text(size = 8), 
    axis.text.x = element_text(angle = 45, hjust = 1),
    legend.position = "left", 
    legend.text = element_text(size = 6), 
    legend.title = element_text(size = 6), 
    legend.margin = margin(t = 5, r = 10, b = 0, l = 10) 
  )

```

Across the 19-year span, the data for states which have not implemented
an abortion policy showcases a wide range of trends. Several states
demonstrate a peak in birth rates at various points during the study
period. While there is variation between the states, the overall
trajectory indicated that the early to late 2000s were a period of
higher birth rates, with a shift toward lower birth rates post 2010. It
is important to note that this data in Figure 3 showcases the relative
average number of births per 1000 people and fluctuates from 10 to
approximately 16.

```{r echo=FALSE, message=FALSE, results='asis'}
#| label: fig-4
# Define file paths 
policy_data_path <- "../data/analysis_data/annual_policy/policyvars01_19.dta"
population_data_path <- "../data/UKCPR_National_Welfare_Data_Update_020623.xlsx"
births_data_path <- "../data/analysis_data/annual_policy/numbirths_2001_2019.csv"

# Load and prepare the abortion policy data
abortion_policy <- read_dta(policy_data_path) %>%
  filter(year >= 2001, year <= 2019) %>%
  select(year, stname, delay)

# Load and prepare the population data
population_data <- read_excel(population_data_path, sheet = "Data") %>%
  select(state_name, year, Population) %>%
  filter(year >= 2001, year <= 2019) %>%
  rename(stname = state_name)

# Load and prepare the number of births data
numbirths_data <- read_csv(births_data_path) %>%
  rename(births = numbirth1544) %>%
  filter(year >= 2001, year <= 2019)

# Merge the datasets
births_population_data <- numbirths_data %>%
  left_join(population_data, by = c("year", "stname")) %>%
  mutate(relative_births = (births / Population) * 1000)

combined_data <- abortion_policy %>%
  left_join(births_population_data, by = c("year", "stname"))

# Define the groups based on predefined 'states_no_delay' and 'states_with_delay'
states_no_delay <- c("AK", "AL", "CA", "CO", "CT", "DC", "DE", "FL", "HI", "IA", "IL", "MA", "MD", "ME", "MT", "NH", "NJ", "NM", "NV", "NY", "OR", "RI", "TN", "VT", "WA", "WY")
states_with_delay <- c("AR", "ID", "IN", "KS", "KY", "LA", "MI", "MO", "MS", "ND", "NE", "OH", "PA", "SC", "SD", "UT", "VA", "WI")

# Filter data for states with delay and plot
combined_data_with_delay <- combined_data %>%
  filter(stname %in% states_with_delay)

ggplot(data = combined_data_with_delay, aes(x = year, y = relative_births, group = stname, color = stname)) +
  geom_line() +
  labs(title = "Relative Average Number of Births in States With Delay (2001-2019)",
       x = "Year",
       y = "Relative Average Number of Births per 1000 People") +
  theme_minimal() +
  theme(
    plot.title = element_text(size = 10),  
    axis.title.y = element_text(size = 10),  
    axis.text.x = element_text(angle = 45, hjust = 1),  
    legend.position = "right",  
    legend.text = element_text(size = 8),  
    legend.title = element_text(size = 10),  
    legend.key.size = unit(0.1, 'cm'),  
    legend.spacing.y = unit(0.1, 'cm')  
  )
```

In contrast, figure 4 looks at the trend of relative average number of
births which have consistently implemented a delay policy. Similar to
states which have not implemented a delay policy, birth rates for these
states have generally decreased over the 19 year period. Over this
period, the relative average number of births per 1000 people fluctuated
significantly, ranging from around 10 to over 20.

```{r echo=FALSE, message=FALSE, results='asis'}
#| label: fig-5

# Define file paths 
policy_data_path <- "../data/analysis_data/annual_policy/policyvars01_19.dta"
population_data_path <- "../data/UKCPR_National_Welfare_Data_Update_020623.xlsx"
births_data_path <- "../data/analysis_data/annual_policy/numbirths_2001_2019.csv"

# Load and prepare the abortion policy data
abortion_policy <- read_dta(policy_data_path) %>%
  filter(year >= 2001, year <= 2019) %>%
  select(year, stname, delay)

# Load and prepare the population data
population_data <- read_excel(population_data_path, sheet = "Data") %>%
  select(state_name, year, Population) %>%
  filter(year >= 2001, year <= 2019) %>%
  rename(stname = state_name)

# Load and prepare the number of births data
numbirths_data <- read_csv(births_data_path) %>%
  rename(births = numbirth1544) %>%
  filter(year >= 2001, year <= 2019)

# Merge the datasets
births_population_data <- numbirths_data %>%
  left_join(population_data, by = c("year", "stname")) %>%
  mutate(relative_births = (births / Population) * 1000)

combined_data <- abortion_policy %>%
  left_join(births_population_data, by = c("year", "stname"))

# Define the groups based on predefined 'states_no_delay' and 'states_with_delay'
states_no_delay <- c("AK", "AL", "CA", "CO", "CT", "DC", "DE", "FL", "HI", "IA", "IL", "MA", "MD", "ME", "MT", "NH", "NJ", "NM", "NV", "NY", "OR", "RI", "TN", "VT", "WA", "WY")
states_with_delay <- c("AR", "ID", "IN", "KS", "KY", "LA", "MI", "MO", "MS", "ND", "NE", "OH", "PA", "SC", "SD", "UT", "VA", "WI")

# States which had a change in policy and plot
policy_changes <- data.frame(
  stname = c("AZ", "GA", "MN", "NC", "OK", "TX", "WV"),
  year_implemented = c(2009, 2005, 2003, 2011, 2005, 2003, 2003)
)

transition_states_data <- combined_data %>%
  filter(stname %in% policy_changes$stname)

# Filter and plot for states which had a change in policy
transition_states_data <- combined_data %>%
  filter(stname %in% policy_changes$stname)

ggplot(data = transition_states_data, aes(x = year, y = relative_births, group = stname, color = stname)) +
  geom_line() +
  geom_vline(data = policy_changes, aes(xintercept = year_implemented, color = stname), linetype="dashed") +
  geom_text(data = policy_changes, aes(x = year_implemented, y = Inf, label = year_implemented), vjust = -0.5, color = "black") +
  scale_x_continuous(breaks = seq(2001, 2019, by = 2)) +
                       labs(title = "Transition to Policy Delay and Relative Average Number of Births (2001-2019)",
                            x = "Year",
                            y = "Relative Average Number of Births per 1000 People",
                            color = "State") +
                       theme_minimal() +
                       theme(
                         plot.title = element_text(size = 10),
                         axis.title.y = element_text(size = 10),
                         axis.text.x = element_text(angle = 45, hjust = 1)
                         )
```

Lastly, figure 5 looks at states which transitioned from having no
abortion delay to policy to implementing this policy at various years
(indicated by the broken lines). A common trend among these states was
an initial period of relatively higher birth rates in the early 2000s,
with subsequent declines over the years. This trend also closely follows
what observed for states which implemented an abortion delay policy and
those which have not implemented such a policy.

### Difference-in-Difference

```{r echo=FALSE, message=FALSE, results='asis'}
#| label: fig-11
# Define file paths 
policy_data_path <- "../data/analysis_data/annual_policy/policyvars01_19.dta"
population_data_path <- "../data/UKCPR_National_Welfare_Data_Update_020623.xlsx"
births_data_path <- "../data/analysis_data/annual_policy/numbirths_2001_2019.csv"

# Load and prepare the abortion policy data
abortion_policy <- read_dta(policy_data_path) %>%
  filter(year >= 2001, year <= 2019) %>%
  select(year, stname, delay)

# Load and prepare the population data
population_data <- read_excel(population_data_path, sheet = "Data") %>%
  select(state_name, year, Population) %>%
  filter(year >= 2001, year <= 2019) %>%
  rename(stname = state_name)

# Load and prepare the number of births data
numbirths_data <- read_csv(births_data_path) %>%
  rename(births = numbirth1544) %>%
  filter(year >= 2001, year <= 2019)

# Merge the datasets
births_population_data <- numbirths_data %>%
  left_join(population_data, by = c("year", "stname")) %>%
  mutate(relative_births = (births / Population) * 1000)

combined_data <- abortion_policy %>%
  left_join(births_population_data, by = c("year", "stname"))

# Difference-in-Difference Analysis for NC with a policy change in 2011 compared to a no-delay state
did_data <- combined_data %>%
  filter(stname %in% c("NC", "TN"), year >= 2001, year <= 2019) %>%
  mutate(treatment = if_else(stname == "NC", 1, 0),
         post_policy = if_else(year >= 2011, 1, 0),
         treatment_x_post = treatment * post_policy)

# Fit the DiD model
did_model <- lm(relative_births ~ treatment + post_policy + treatment_x_post, data = did_data)

# Summary of the DiD model and robust standard errors
summary_did <- summary(did_model)
robust_summary <- coeftest(did_model, vcov = vcovHC(did_model, type = "HC1"))

# Extract coefficients and their standard errors
coefficients <- coef(robust_summary)
std_errors <- sqrt(diag(vcovHC(did_model, type = "HC1")))

# Combine coefficients and standard errors into a data frame
coef_df <- data.frame(Estimate = coefficients, `Std. Error` = std_errors)

# Output the robust summary as a table
kable(coef_df, format = "markdown", caption = "Table 1")

# Plotting for the DiD analysis
ggplot(did_data, aes(x = year, y = relative_births, color = stname)) +
  geom_line() +
  geom_vline(xintercept = 2011, linetype="dashed", color = "red") + # Policy change year
  labs(title = "Relative Births in NC (Treatment) and TN (Comparison) Over Time",
       x = "Year", y = "Relative Births per 1000 People") +
  theme_minimal() +
                       theme(
                         plot.title = element_text(size = 10)
                        )

```

These abortion delay policies created natural experiments which can be
used to apply a Difference-in-Differences approach. The results from
comparing North Carolina (treatment) with Tennessee (control) are as
follow: - The intercept ($B_0$) estimate is 13.4345, suggesting that the
baseline relative number of births per 1000 people is 13.4345 in the
absence of treatment and before the policy change. - The treatment
effect ($B_1$) is estimated at 0.4887, which is statistically
significant (p = 0.01078). This indicates that, prior to the policy
change, NC had 0.4887 more births per 1000 people than TN. - The
post_policy effect ($B_2$) is -1.2372, also statistically significant (p
= 1.26e-07), suggesting a general decrease in relative births per 1000
people after 2011 in both states. - The interaction term ($B_3$), which
represents the estimated effect of the policy change, is -0.7770, with a
p-value of 0.00569, indicating that the policy delay in NC is associated
with a reduction of 0.7770 births per 1000 people relative to TN after
the policy implementation. - The model explains a significant portion of
the variance in relative births per 1000 people ($R\^2$ = 0.8267),
indicating a good fit.

## Education

```{r echo=FALSE, message=FALSE}
#| label: fig-6

#Education

# Load Education data
educ_data <- read_csv("../data/analysis_data/educ/numbirths_educ_2044.csv")

# Creating the plot
ggplot(educ_data, aes(x = year)) +
  geom_line(aes(y = numbirth_lesshs, color = "Less than HS"), linewidth = 1) +
  geom_line(aes(y = numbirth_hsgrad, color = "HS Graduate"), linewidth = 1) +
  geom_line(aes(y = numbirth_somecoll, color = "Some College"), linewidth = 1) +
  geom_line(aes(y = numbirth_colgrad, color = "College Graduate"), linewidth = 1) +
  geom_line(
    aes(y = numbirthalleduc, color = "All Education Levels"),
    linewidth = 1,
    linetype = "dashed"
  ) +
  scale_color_manual(
    values = c(
      "Less than HS" = "blue",
      "HS Graduate" = "green",
      "Some College" = "orange",
      "College Graduate" = "red",
      "All Education Levels" = "grey"
      )
  ) +
  labs(
    title = "Number of Births by Education Level Over Time",
    x = "Year",
    y = "Number of Births",
    color = "Education Level"
  ) +
  theme_minimal() +
  theme(legend.title = element_text(size = 12),
        legend.text = element_text(size = 10))

# Display the plot
ggsave("births_by_education_level_over_time.png",
       width = 10,
       height = 6)

```

The figure 6 displays the trend in number of births segmented by education achievement from 1990 to 2019. The education levels all follow a similar trend with the exception of less than HS. For this education level, there is a decline in number of births post 2008 whereas the other education levels showed an increase in birth rates.

### State Political Affiliation and Birth Rates

```{r echo=FALSE, message=FALSE}
#| label: fig-7
# https://stackoverflow.com/questions/29381652/plotting-us-heat-map-with-text

#reading the UKNCPR Datat
poldata <- read_csv(here("data", "analysis_data", "UKCPR_National_Welfare_Data_Update_020623.csv"))

poldata <- poldata %>%
  select(state_name,
         year,
         Population,
         `Fraction of State Senate that is Democrat`)

# Create a mapping of state abbreviations to full names
state_abbrevs <- state.abb
state_names <- state.name

# Create a data frame for the mapping
state_mapping <-
  data.frame(abbrev = tolower(state_abbrevs),
             name = tolower(state_names))

#Get numbirths data
numbirths_data <- read_csv(here("data", "analysis_data", "annual_policy", "numbirths_2001_2019.csv"))
births_2019 <- numbirths_data %>% filter(year == 2019)

# Convert state abbreviations to full names in your datasets
numbirths_2019_fullname <- births_2019 %>%
  mutate(stname = tolower(stname)) %>%
  left_join(state_mapping, by = c("stname" = "abbrev")) %>%
  select(-stname) %>%
  rename(stname = name)

poldata_2019 <- poldata %>% filter(year == 2019)

poldata_2019_fullname <- poldata_2019 %>%
  mutate(state_name = tolower(state_name)) %>%
  left_join(state_mapping, by = c("state_name" = "abbrev")) %>%
  select(-state_name) %>%
  rename(state_name = name)

# Merge poldata with numbirths_data using full state names
data_merged <-
  left_join(numbirths_2019_fullname,
            poldata_2019_fullname,
            by = c("stname" = "state_name"))

states_map <- map_data("state")

# Merge with states map data
map_data_merged <-
  left_join(states_map, data_merged, by = c("region" = "stname"))

# Plot
ggplot(map_data_merged) +
  geom_polygon(
    aes(
      x = long,
      y = lat,
      group = group,
      fill = `Fraction of State Senate that is Democrat`
    ),
    color = "white"
  ) +
  geom_text(
    data = map_data_merged %>% group_by(region) %>% summarise(
      long = mean(long),
      lat = mean(lat),
      numbirths = first(numbirth1544)
    ),
    aes(x = long, y = lat, label = numbirths),
    size = 3,
    check_overlap = TRUE
  ) +
  scale_fill_gradient(low = "red",
                      high = "blue",
                      name = "Democrat Fraction") +
  coord_fixed(1.3) +
  labs(title = "2019 Births and Democrat Fraction in State Senate by State") +
  theme_void() +
  theme(legend.position = "right")
```

TODO: Results description for Political Affiliation

## Child Support Enforcement

```{r echo=FALSE, message=FALSE}
child_enforce_data <- read_dta("../data/raw_data/annual_policy/policyvars01_19.dta")
```

The number of people with ACA Medicaid increased dramatically from 2000
to about 2014. As figure 8 depicts, since 2014, the number of people
with Medicaid has slowly increased non-decreasingly till 2020. The ACA
Medicaid expansion has been adopted and implemented by states at
different times, which explains some of the dramatic increases within
the data.

```{r echo=FALSE, message=FALSE, fig.cap= "Expansion of Medicaid from 2000 to 2020"}
#| label: fig-8
#| 
# medicaid benefits
ggplot(child_enforce_data, aes(x = year, y = aca_mcaid_exp)) +
  geom_col(fill = "#1f77b4" ) + labs(x = "Year", y = "Number of people with ACA Medicaid ", title = "People with ACA Medicaid Over Time (Years)") + theme_minimal() +  theme(plot.title = element_text(hjust = 0.5))  
```

In Figure 9, there has been a linear decrease of the maximum welfare
benefits from 2000 to 2015. From 2015 to 2020 there has been a small
rise in the maximum welfare benefits. Overall, there has been a loss in
the maximum welfare benefits, despite the small rise.

```{r echo=FALSE, message=FALSE}
#| label: fig-9
# max well fair benefits
ggplot(child_enforce_data, aes(x=year, y=logben3)) +
    geom_point()  + labs(x = "Year", y = "Max Welfare Benefits", title = "Max Welfare Benefits Over Time (Years)") + theme_minimal() +  theme(plot.title = element_text(hjust = 0.5))  
```

In Figure 10, the blue portion of each column represents the number of
parents who have consent for an abortion. The red portion represents the
number of parents who don’t have consent for an abortion. There is a
slight increase in consent for abortion from 2001 to 2019.

```{r echo=FALSE, message=FALSE}
 #| label: fig-10
# Parental
ggplot(child_enforce_data, aes(x = factor(year), fill = factor(parental))) +
  geom_bar(position = "stack") +
  labs(x = "Year", y = "Parental Consent for Abortion", fill = "Interest", title = "Distribution of Interest in Abortion by Parent Over Time (Years)") + theme_minimal() +  # Apply minimal theme
  theme(plot.title = element_text(hjust = 0.5), axis.text.x = element_text(angle = 45, vjust = 0.5))  
```

# Discussion:

## Abortion Policy

Through the difference-in-difference analysis, the results indicate a
decrease of approximately 0.777 births per 1000 people in North Carolina
post policy implementation relative to TN, which did not implement the
abortion delay policy. This decrease is statistically significant,
suggesting that implementing an abortion delay policy may have a real
effect on birth rates.

The observed decrease in birth rates in North Carolina post 2011 could
be partially explained by increased barriers to access as highlighted by
de Londras et al. whose review supports the results derived in this
paper. This review summarizes the effects of mandated waiting periods
(MWPs) on outcomes connected to abortion. Their analysis suggests that
MWPs may cause delays in obtaining abortion services, which may lower
the number of births particularly in young, unmarried women. Noting the
disproportionate impact on marginalized populations, such as teenagers,
persons of color, and those with lower socioeconomic standing, they also
draw attention to the broader human rights concerns of MWPs.

The results of the DiD found in this paper are supported by strong
standard errors. However, we need to consider the limitations of the DiD
approach as there are several other factors which may impact birth rates
such as economic fluctuations and access to healthcare. Future
iterations of this paper could expand on this research by conducting
multi-state analysis for a variety of confounding factors.

## Political affiliation and birth rates

From our analysis of political affiliations over the years, there are
some relation between the two, especially in years following aggresive
rhetorics or policies about reproductive rights. This indicates that
periods following aggressive political rhetoric or significant policy
changes regarding reproductive rights often coincide with noticeable
changes in birth rates. This trend is observable across various states
and is particularly pronounced in areas where political affiliation (#
of democrats in state senate) change very differently from one side to
the other. The rhetoric employed by political leaders and the nature of
the policies they implement can have a substantial impact on
individuals' choices regarding family planning. For instance, policies
that limit access to reproductive healthcare, including contraception
and abortion services, can lead to an increase in birth rates,
especially in regions where alternative options are
scarce.[@Nargund2009DecliningBirthRate]

Conversely, political movements that advocate for and expand access to
reproductive healthcare are often associated with a stabilization or
decrease in birth rates. This can be attributed to increased autonomy
over reproductive choices, allowing individuals to make informed
decisions about if and when to have children.

## Education
We observe a convergence in birth rates across different educational categories based on the data for number of births by education level. This could be a reflection of larger societal shifts in family planning preferences and educational achievement. College graduates' stable fertility rate followed by a fall in births could mean that this group is delaying or reducing the size of their family, possibly for financial or career-related reasons. The overall decline in births over the past few years across all educational levels may reflect a general trend toward smaller families, or it may be the result of other outside factors not shown in the graph, such as shifting social norms, shifting economic conditions, or expanding access to reproductive healthcare.

## Child Support Enforcement

(ACA)’s Medicaid is “intended to address systemic health inequalities
for millions of Americans who lack health insurance” Medicaid is
“expected to provide coverage to low-income individuals,” and also
covers low-income pregnant women. Excluding the states where abortion is
banned, Medicaid covers abortions, making it more accessible. According
to figure 8, More people continue to have medicaid as seen based on the
trend from 2000 to 2020, making an affordable abortion more accessible.
In figure 10, we can see that there is an increase in parental consent
for abortion. In other words, more parents are increasingly interested
in getting an abortion every year. A potential motivator for parents to
want an abortion is that as a low-income family, it is harder to take
care of a larger family. Welfare is “federal government-sponsored
assistance programs for individuals and families in need.” In figure 9,
it is evident that there is a decrease in maximum welfare benefits,
making it less affordable for lower-income families to have a larger
family size.The increasing demand for an abortion and the insurance
covering it may contribute to a decline in birthrates.

## 4.2 Weakness and Next Steps

An improvement we can add to this paper is adding additional economical
context along with our data such as investigating events like the
Recession that took place in 2008, and the conflict regarding
reproductive right policies. Without providing context to current events
in the time period, our paper gives a weaker justification for the
decline in birthrates. Next steps for improving our analysis include not
only providing context for the events happening in the US, but focusing
on the birth rates on a more micro level. Since each state within the US
has its own laws, it will provide more insight to focus on the birth
rates within each state. Additionally, to improve our analysis
furthermore, we can compare the US to other countries that are similar
from an economical standpoint and compare their birth rate to the US.

\newpage

# References
@Dataset{ukcprWelfareData,
  title = {UKCPR National Welfare Data, 1980-2021},
  author = {{University of Kentucky Center for Poverty Research}},
  year = {2023},
  month = {Feb},
  address = {Lexington, KY},
  url = {http://ukcpr.org/resources/national-welfare-data%7D,
  note = {Accessed: <Feb 15, 2024>},
}

@Manual{citeR,
    title = {R: A Language and Environment for Statistical Computing},
    author = {{R Core Team}},
    organization = {R Foundation for Statistical Computing},
    address = {Vienna, Austria},
    year = {2023},
    url = {https://www.R-project.org/},
  }

@InBook{Grundy2015Demography,
  author    = {Emily Grundy and Michael Murphy},
  editor    = {Roger Detels and Martin Gulliford and Quarraisha Abdool Karim and Chorh Chuan Tan},
  title     = {Demography and public health},
  booktitle = {Oxford Textbook of Global Public Health},
  publisher = {Oxford University Press},
  address   = {Oxford, England},
  edition   = {6},
  pages     = {718--735},
  year      = {2015},
}

@Article{Kearney2022Puzzle,
  author    = {Melissa S. Kearney and Phillip B. Levine and Luke Pardue},
  title     = {The Puzzle of Falling US Birth Rates since the Great Recession},
  journal   = {Journal of Economic Perspectives},
  year      = {2022},
  volume    = {36},
  number    = {1},
  pages     = {151--176},
  doi       = {10.1257/jep.36.1.151},
}

@Article{thereferencecanbewhatever,
    title = {Welcome to the {tidyverse}},
    author = {Hadley Wickham and Mara Averick and Jennifer Bryan and Winston Chang and Lucy D'Agostino McGowan and Romain François and Garrett Grolemund and Alex Hayes and Lionel Henry and Jim Hester and Max Kuhn and Thomas Lin Pedersen and Evan Miller and Stephan Milton Bache and Kirill Müller and Jeroen Ooms and David Robinson and Dana Paige Seidel and Vitalie Spinu and Kohske Takahashi and Davis Vaughan and Claus Wilke and Kara Woo and Hiroaki Yutani},
    year = {2019},
    journal = {Journal of Open Source Software},
    volume = {4},
    number = {43},
    pages = {1686},
    doi = {10.21105/joss.01686},
  }

  @Article{Nargund2009DecliningBirthRate,
    title = {Declining birth rate in Developed Countries: A radical policy re-think is required},
    author = {Nargund, G.},
    year = {2009},
    journal = {Facts Views Vis Obgyn},
    volume = {1},
    number = {3},
    pages = {191-193},
    PMID = {25489464},
    PMCID = {PMC4255510},
}

@Manual{citeLahman,
    title = {Lahman: Sean `Lahman' Baseball Database},
    author = {Michael Friendly and Chris Dalzell and Martin Monkman and Dennis Murphy},
    year = {2020},
    note = {R package version 8.0-0},
    url = {https://CRAN.R-project.org/package=Lahman},
  }

@article{gebru2021datasheets,
  title={Datasheets for datasets},
  author={Gebru, Timnit and Morgenstern, Jamie and Vecchione, Briana and Vaughan, Jennifer Wortman and Wallach, Hanna and Iii, Hal Daum{\'e} and Crawford, Kate},
  journal={Communications of the ACM},
  volume={64},
  number={12},
  pages={86--92},
  year={2021},
  publisher={ACM New York, NY, USA}
}

@Misc{rstanarm,
    title = {rstanarm: {Bayesian} applied regression modeling via {Stan}},
    author = {Ben Goodrich and Jonah Gabry and Imad Ali and Sam Brilleman},
    note = {R package version 2.21.3},
    year = {2022},
    url = {https://mc-stan.org/rstanarm/},
  }

@Manual{palmerpenguins,
    title = {palmerpenguins: Palmer Archipelago (Antarctica) penguin data},
    author = {Allison Marie Horst and Alison Presmanes Hill and Kristen B Gorman},
    year = {2020},
    note = {R package version 0.1.0},
    doi = {10.5281/zenodo.3960218},
    url = {https://allisonhorst.github.io/palmerpenguins/},
  }

@Article{rohan,
    title = {Welcome to the {tidyverse}},
    author = {Hadley Wickham and Mara Averick and Jennifer Bryan and Winston Chang and Lucy D'Agostino McGowan and Romain François and Garrett Grolemund and Alex Hayes and Lionel Henry and Jim Hester and Max Kuhn and Thomas Lin Pedersen and Evan Miller and Stephan Milton Bache and Kirill Müller and Jeroen Ooms and David Robinson and Dana Paige Seidel and Vitalie Spinu and Kohske Takahashi and Davis Vaughan and Claus Wilke and Kara Woo and Hiroaki Yutani},
    year = {2019},
    journal = {Journal of Open Source Software},
    volume = {4},
    number = {43},
    pages = {1686},
    doi = {10.21105/joss.01686},
  }

@Manual{,
  title = {readxl: Read Excel Files},
  author = {Hadley Wickham and Jennifer Bryan},
  year = {2023},
  note = {https://readxl.tidyverse.org, https://github.com/tidyverse/readxl},
}

@Article{,
    title = {Welcome to the {tidyverse}},
    author = {Hadley Wickham and Mara Averick and Jennifer Bryan and
      Winston Chang and Lucy D'Agostino McGowan and Romain François and
      Garrett Grolemund and Alex Hayes and Lionel Henry and Jim Hester
      and Max Kuhn and Thomas Lin Pedersen and Evan Miller and Stephan
      Milton Bache and Kirill Müller and Jeroen Ooms and David Robinson
      and Dana Paige Seidel and Vitalie Spinu and Kohske Takahashi and
      Davis Vaughan and Claus Wilke and Kara Woo and Hiroaki Yutani},
    year = {2019},
    journal = {Journal of Open Source Software},
    volume = {4},
    number = {43},
    pages = {1686},
    doi = {10.21105/joss.01686},
  }

@Article{,
    title = {Diagnostic Checking in Regression Relationships},
    author = {Achim Zeileis and Torsten Hothorn},
    journal = {R News},
    year = {2002},
    volume = {2},
    number = {3},
    pages = {7--10},
    url = {https://CRAN.R-project.org/doc/Rnews/},
  }

@Manual{sandwich,
  title = {sandwich: Robust Covariance Matrix Estimators},
  author = {Achim Zeileis},
  year = {2022},
  note = {R package version 3.0-1},
  url = {https://CRAN.R-project.org/package=sandwich},
}

@Article{,
    title = {Econometric Computing with {HC} and {HAC} Covariance
      Matrix Estimators},
    author = {Achim Zeileis},
    journal = {Journal of Statistical Software},
    year = {2004},
    volume = {11},
    number = {10},
    pages = {1--17},
    doi = {10.18637/jss.v011.i10},
  }

@Article{,
    author = {Hadley Wickham},
    title = {testthat: Get Started with Testing},
    journal = {The R Journal},
    year = {2011},
    volume = {3},
    pages = {5--10},
    url =
      {https://journal.r-project.org/archive/2011-1/RJournal_2011-1_Wickham.pdf},
  }

@Article{delondras2022impact,
  title = {The impact of mandatory waiting periods on abortion-related outcomes: a synthesis of legal and health evidence},
  author = {Fiona de Londras and Amanda Cleeve and Maria I. Rodriguez and Alana Farrell and Magdalena Furgalska and Antonella Lavelanet},
  journal = {BMC Public Health},
  volume = {22},
  pages = {1232},
  doi = {10.1186/s12889-022-13620-z},
  url = {https://bmcpublichealth.biomedcentral.com/articles/10.1186/s12889-022-13620-z},
}